
Which agencies do you want to pull data from?
```{r Define directories and choose data source}
rm(list = ls())

library(ggplot2)
library(gridExtra)

baseDir <- 'C:/KPONEIL/'

dataInDir <- paste0(baseDir, 'GitHub/projects/temperatureProject/dataIn/')
dataOutDir <- paste0(baseDir, 'GitHub/projects/temperatureProject/dataOut/')
graphsDir <- paste0(baseDir, 'GitHub/projects/temperatureProject/graphs/')

CTDEP  <- T
MAFW   <- T
MAUSGS <- T
NHFG   <- T
NHDES  <- T

USFS   <- T
VTFWS  <- T

MEDMR  <- F
MTUSGSYellowstone <- F
MTUSGSGlacier <- F

sourceChoice <- list( CTDEP,   MAFW,   MAUSGS,   NHFG,   NHDES,   MEDMR,   USFS,   VTFWS,    MTUSGSYellowstone,   MTUSGSGlacier)
sourceNames  <- c   ('CTDEP', 'MAFW', 'MAUSGS', 'NHFG', 'NHDES', 'MEDMR', 'USFS', 'VTFWS',  'MTUSGSYellowstone', 'MTUSGSGlacier')

dataSource <- sourceNames[sourceChoice == T]

sourceList <- paste0(paste0(dataSource, collapse = '_'))#, '_', Sys.Date())

#global vars}
dpiIn <- 400
```


```{r load data}

Run <- "_Run1"

#Load "et" for the agencies:
load(paste0(dataOutDir, sourceList,  Run, '/et_', sourceList, '.RData'))

#These are duplicates that get removed from daymet dataframe before merging:
removeColsCovariate <- c('Latitude', 'Longitude')
#removeColsMaster    <- c('AgencyID', 'agency', 'temp', 'airTemp', 'Latitude', 'Longitude')

#Load in covariate data to merge into slopes df [no day data]
for ( i in 1:length(dataSource)){

  #Load covariate data to be merged into slopes df [no day data]
  load(paste0(dataInDir, dataSource[i], '/covariateData_', dataSource[i], '.RData')) #Jan 2014 covariateData [fixed over time] 
  covariateData <- covariateData[,-which(names(covariateData) %in% removeColsCovariate)]
  if ( i == 1) {covs <- covariateData} else (covs <- rbind(covs, covariateData))
  
  ##Load daymet climate data to be merged into et:
  #load(paste0(dataInDir, dataSource[i], '/streamTempSitesObservedClimateData_', dataSource[i], '.RData')) 
  #masterData <- masterData[, -which(names(masterData) %in% removeColsMaster)]
  #if ( i == 1) {newDay <- masterData} else ( rbind(newDay, masterData) )
  
} 
  
#masterData    <- newDay
covariateData <- covs

#et <- merge(et, masterData, by = c('site', 'year', 'dOY'), all.x=T, sort = F )


#Overwrite NaNs with NAs:
covariateData <- replace(covariateData, is.na(covariateData), NA)

#covariateData$site <- covariateData$Site
covariateData$site <- as.character(covariateData$site)

et <- merge(et, covariateData, by = 'site', all.x=T, sort = F )


```


```{r Make some helpful plots}

subGraphsDir  <- paste0(graphsDir,  sourceList)

#Histogram of rising and falling limbs of segmented regression used in summer BP estimation.
summerImpound  <- ggplot( et ) + 
  geom_point(aes(PercentImpoundedOpen, summerBP  ), colour='red' ) +
   #  ylim(c(-10,10))+
   #  xlim(c(80,120))+
    theme(strip.text.x = element_text(size=6)) +
    ggtitle('SummerBP vs PercentImpounded')

fallImpound  <- ggplot( et ) + 
  geom_point(aes(PercentImpoundedOpen, fallBP  ), colour='red' ) +
   #  ylim(c(-10,10))+
   #  xlim(c(80,120))+
    theme(strip.text.x = element_text(size=6)) +
    ggtitle('FallBP vs PercentImpounded')

gImpound  <- arrangeGrob( summerImpound,  fallImpound,  ncol=1 )
ggsave(plot=gImpound, file=paste0(subGraphsDir,'/PcntImpndBP.png'),dpi=300,width=6,height=8, units='in', scale=2)


summerDamCount  <- ggplot( et ) + 
  geom_point(aes(TNC_DamCount, summerBP  ), colour='red' ) +
   #  ylim(c(-10,10))+
     xlim(c(0,100))+
    theme(strip.text.x = element_text(size=6)) +
    ggtitle('SummerBP vs TNC_DamCount')

fallDamCount  <- ggplot( et ) + 
  geom_point(aes(TNC_DamCount, fallBP  ), colour='red' ) +
   #  ylim(c(0,100))+
     xlim(c(0,100))+
    theme(strip.text.x = element_text(size=6)) +
    ggtitle('FallBP vs TNC_DamCount')


gDamCount <- arrangeGrob( summerDamCount, fallDamCount, ncol=1 )
ggsave(plot=gDamCount,file=paste0(subGraphsDir,'/DamCountBP.png'), dpi=300,width=6,height=8, units='in', scale=2)




risingTempPctImp  <- ggplot( et ) + 
  geom_point(aes(PercentImpoundedOpen, risingTemp  ), colour='red' ) +
   #  ylim(c(-10,10))+
     xlim(c(0,100))+
    theme(strip.text.x = element_text(size=6)) +
    ggtitle('Rising Temp Slope (temp/dOY) vs PercentImpoundedOpen')

fallingTempPctImp  <- ggplot( et ) + 
  geom_point(aes(PercentImpoundedOpen, fallingTemp  ), colour='red' ) +
   #  ylim(c(0,100))+
     xlim(c(0,100))+
    theme(strip.text.x = element_text(size=6)) +
    ggtitle('Falling Temp Slope (temp/dOY) vs PercentImpoundedOpen')


gDamCount <- arrangeGrob( risingTempPctImp, fallingTempPctImp, ncol=1 )
ggsave(plot=gDamCount,file=paste0(subGraphsDir,'/riseFallPctImp.png'), dpi=300,width=6,height=8, units='in', scale=2)




```

  
  
  

  
  
  
  
  
  
  
  
  
  
  theme(strip.text.x = element_text(size=6)) +
  scale_x_continuous(breaks = c(seq(from = -2, to = 2, by = .2)),
                     labels = c(seq(from = -2, to = 2, by = .2))) +
  ggtitle('Rising limb of seg-regression for SummerBP')

gFallingTemp <- ggplot( breaks) + 
  geom_histogram(aes(fallingTemp), binwidth = .01) + 
  theme(strip.text.x = element_text(size=6)) +
  scale_x_continuous(breaks = c(seq(from = -3, to = 2, by = .2)),
                     labels = c(seq(from = -3, to = 2, by = .2))) +
  ggtitle('Falling limb of seg-regression for SummerBP')

gHist <- arrangeGrob( gRisingTemp, gFallingTemp, ncol=1)

gBP <- ggplot( temp, aes(dOY,tempIndex)) + 
    geom_point(aes(dOY,movingSD  ),size = 0.5, colour='red' ) +
    geom_point(aes(dOY,movingMean),size = 0.5, colour='blue') +
    geom_point(size=0.5) +
    geom_hline( aes(yintercept=as.numeric(quantileLo)), colour='black') +
    geom_hline( aes(yintercept=as.numeric(quantileHi)), colour='black') +
    geom_vline( aes(xintercept=as.numeric(springBP  )), colour='green') +
    geom_vline( aes(xintercept=as.numeric(fallBP    )), colour='orange')+
    geom_vline( aes(xintercept=as.numeric(summerBP  )), colour='blue')  +
    ylim(c(-10,10))+
   #  xlim(c(80,120))+
    theme(strip.text.x = element_text(size=6))+
    ggtitle(paste(temp$site[1],sep=" ")) +
    facet_wrap(~year)
  

```







