# this file is for multiple sites, also works for a single site

# use of max vs min airTemp in addition to mean air temp

```{r load libraries}
rm(list=ls())

library(ggplot2)
library(GGally)
library(gridExtra)
library(reshape2)
library(mgcv)
library(nlme)
library(plyr)
library(segmented)
library(zoo)
library(ggmap)
library(pls)
library(MASS)
library(lme4)
library(DataCombine) # for the slide function

#baseDir <- 'C:/KPONEIL/GitHub/projects/temperatureProject/'  
# Kyle
#baseDir <- '/Users/Dan/Documents/Research/Stream_Climate_Change/temperatureProject/'  # Dan
#baseDir <- 'D:/GitHub/projects/temperatureProject/'                                   # Ben
# UNH
baseDir <- 'C:/Users/dhocking/Documents/temperatureProject/'

dataInDir <- paste0(baseDir, 'dataIn/')
dataOutDir <- paste0(baseDir, 'dataOut/')
graphsDir <- paste0(baseDir, 'graphs/')

source(paste0(baseDir, 'code/functions/temperatureModelingFunctions.R'))
setwd(baseDir)
```

Which agencies do you want to pull data from?
```{r Choose data source}

#Northeast
CTDEP  <- T
MAFW   <- T
MAUSGS <- T
MADEP <- T #MADEEP?
NHFG   <- T
NHDES  <- T
USFS   <- T
VTFWS  <- T
MEDMR  <- T

#Montana
MTUSGSYellowstone <- F
MTUSGSGlacier <- F

sourceChoice <- list( CTDEP,   MAFW,   MAUSGS, MADEP,   NHFG,   NHDES,   MEDMR,   USFS,   VTFWS,    MTUSGSYellowstone,   MTUSGSGlacier)
sourceNames  <- c   ('CTDEP', 'MAFW', 'MAUSGS', 'MADEP', 'NHFG', 'NHDES', 'MEDMR', 'USFS', 'VTFWS',  'MTUSGSYellowstone', 'MTUSGSGlacier')
  
dataSource <- sourceNames[sourceChoice == T]

sourceList <- paste0(paste0(dataSource, collapse = '_'), '_', Sys.Date())

setwd(baseDir)
#setwd(paste0(baseDir, 'GitHub/projects/temperatureProject/'))

```

```{r rerun or load previous?}

reRunBP <- T

# Will need to specify date of creation for this to work:
if( !reRunBP ) {
  load(file=paste(dataOutDir, sourceList,'/', 'e_', sourceList, '.RData',sep=''))
  load(file=paste(dataOutDir, sourceList,'/', 'et_', sourceList, '.RData',sep=''))
}

```

```{r Make graphs?}
# settings variables
makeSpringFallGraphs <- F
indexAbnormalSummerGraphs <- F

#Creates folders for graphs if they don't exist:
#-----------------------------------------------
subGraphsDir  <- paste0(graphsDir,  sourceList)
subDataOutDir <- paste0(dataOutDir, sourceList)

if (!file.exists(subGraphsDir)){
  dir.create(file.path(paste0(subGraphsDir)))
  dir.create(file.path(paste0(subGraphsDir, '/summerBP')))
  dir.create(file.path(paste0(subGraphsDir, '/springFallBP')))
  dir.create(file.path(paste0(subGraphsDir, '/problemSites')))
}
  
if (!file.exists(subDataOutDir)){ 
  dir.create(file.path(paste0(subDataOutDir)))
}

```

```{r Load data}
 if ( reRunBP ){
  
  setwd(dataInDir)

  for ( i in 1:length(dataSource)){
    load(paste0(dataSource[i], '/streamTempSitesObservedClimateData_', dataSource[i], '.RData'))  #masterData
    if ( i == 1) {e <- masterData} else ( e <- rbind(e, masterData))
    rm(masterData)
    
    load(paste0(dataSource[i], '/covariateData_', dataSource[i], '.RData'))  #masterData
    if ( i == 1) {d <- covariateData} else ( d <- rbind(d, covariateData))
    rm(covariateData)
  }

d$site <- as.character(d$site)
e <- merge(e,d[,c('site', 'StreamOrder', 'HUC4', 'HUC8', 'HUC12' )], by = 'site', all.x = T, sort = F)

# Needed because R confuses factor/character/numeric. 
e$StreamOrder <- as.character(e$StreamOrder)
e$HUC4       <- as.character(e$HUC4)
e$HUC8       <- as.character(e$HUC8)
e$HUC12      <- as.character(e$HUC12)
}
```

###############################################
# format for input file
# > str(e)
#'data.frame':   622507 obs. of  10 variables:
# $ site     : chr  "CTDEP_1" "CTDEP_1" "CTDEP_1" "CTDEP_1" ...
# $ year     : num  2008 2008 2008 2008 2008 ...
# $ dOY      : num  1 10 100 101 102 103 104 105 106 107 ...
# $ date     : Date, format: "2008-01-01" "2008-01-10" ...
# $ AgencyID : Factor w/ 1479 levels "1","1008","101",..: 1 1 1 1 1 1 1 1 1 1 ...
# $ agency   : chr  "CTDEP" "CTDEP" "CTDEP" "CTDEP" ...
# $ temp     : num  NA NA NA NA NA ...
# $ airTemp  : num  2.25 7.25 8 13.5 12.75 ...
# $ Latitude : num  41.2 41.2 41.2 41.2 41.2 ...
# $ Longitude: num  -73.3 -73.3 -73.3 -73.3 -73.3 ...
##############################################

# need to make sure that dOY is consecutive within a year_site combo, otherwise the slide function will line up non-consecutive dOYs
# If a df doesn't have consec dOY could use functions in zoo or do the indexing explicitly

```{r Data prep}

siteList <- unique(e$site)

#==============================================================================================
# This is the key metric for estimating the synchrony between air and water temp 
e$tempIndex <- (e$temp-e$airTemp)/(e$temp + 0.00000001) # add small # to avoid Inf
#==============================================================================================

# fill in missing dOY so sequence calcs (like moving mean, slope etc) 

# lag by group
e <- e[order(e$site,e$year,e$dOY),]

#For checking the order of e:
#----------------------------
e$count <- 1:length(e$year)

#Get the slope of the air/stream temperature relationship:
#---------------------------------------------------------
e$siteYear <- paste(e$site,e$year,sep='_')
e <- slide(e, Var = "temp", GroupVar = "siteYear", slideBy = +1)
e <- slide(e, Var = "airTemp", GroupVar = "siteYear", slideBy = +1)

e$rise <- e$temp1 - e$temp
e$run  <- e$airTemp1 - e$airTemp
e$slope <- e$rise / e$run
e$length <- sqrt(e$rise^2 + e$run^2)
e$angle <- atan(e$rise/e$run) * 180/pi # need to check radian units...

# for dOY changes
e$waterDelta <-    e$temp1 -    e$temp
e$airDelta   <- e$airTemp1 - e$airTemp

e <- e[order(e$count),]

```

```{r moving mean}
#===============================================================================
#Get moving mean and SD of temp index for each site and put into the data frame:
#===============================================================================
window <- 10 # frame sizefor moving mean, which is centered by default

nSites <- length(siteList)
siteYearCombos <- unique(e[,c('site','year')])
#siteYearCombos$site  <- factor(siteYearCombos$site)

e$movingMean <- NA
e$movingSD <- NA

for (i in 1:nrow(siteYearCombos)){

  print(c(i,as.character(siteYearCombos$site[i]),siteYearCombos$year[i],i/nrow(siteYearCombos)))
  
  currSite <- which(e$site == as.character(siteYearCombos$site[i]) & e$year == siteYearCombos$year[i] )

  #Need this so sites with very short records don't crash the loop.
  if(length(currSite) >= window){currMean <-  rollapply(e$tempIndex[currSite], width=window, fill=NA, mean)} else(currMean <- NA)
  if(length(currSite) >= window){currSD <-    rollapply(e$tempIndex[currSite], width=window, fill=NA, sd)}   else(currSD <- NA)
  
  e$movingMean[currSite] <- currMean
  e$movingSD  [currSite] <- currSD
}

e$meanSDDiff <- e$movingSD - e$movingMean

# just to make sure the merge doens't screw up order
e <- e[order(e$count),]

# look at some raw data
#ggplot( e[e$site == e$site[2] ,], aes(dOY,temp) ) + geom_point() + geom_point(aes(dOY,airTemp),color='red') + facet_grid(site~year)
#table( e$year,e$site,is.na( e$temp ) )
```


```{r Create the breaks dataframe}

breaks <- data.frame(array(NA,c(nrow(siteYearCombos),8)))
names(breaks) <- c('site','year', 'HUC8', 'springBP','summerBP','fallBP','quantileLo','quantileHi')
breaks$fallBPComplete <- FALSE
breaks$springBPComplete <- FALSE
breaks$springOrFallBPComplete <- FALSE
breaks$summerBPComplete <- FALSE

breaks$quantileLo <- NA
breaks$quantileHi <- NA

#Define breakpoint time period and range for tempIndex
beginningDayForCI <- 125
endingDayForCI <- 275
loCI <- 0.001
hiCI <- 0.999

for ( i in 1:nrow(siteYearCombos)){
  print(i)
  breaks$year[i]  <- as.numeric  (siteYearCombos$year[i])
  breaks$site[i]  <- as.character(siteYearCombos$site[i]) #would like this to stay a factor, but it ends up as numeric in 'breaks'. not sure why
  breaks$HUC8[i] <- as.character(d$HUC8[which(d$site == breaks$site[i])])
    
  tmp <- e[e$site == siteYearCombos$site[i] & e$year  %in% siteYearCombos$year[i] & e$dOY %in% beginningDayForCI:endingDayForCI,'tempIndex']
  if (any(!is.na(tmp))){
    TIQ <- quantile(tmp, probs=c(loCI,0.5,hiCI),na.rm=T)
    #  movingSDQ <- quantile(e[e$site == siteList[j] & e$year  %in% year & e$dOY %in% beginningDayForCI:endingDayForCI, 'movingSD'], probs=c(0.005,0.5,0.995),na.rm=T)
    
    breaks$quantileLo[i] <- TIQ[1]
    breaks$quantileHi[i] <- TIQ[3]
  }
}  

```


```{r Spring/fall bp}

#===================================================================================
#Use runs analysis of the movingMean to define spring and fall breakpoints:
#===================================================================================

minCompleteDOYBP1 <- 15
maxCompleteDOYBP1 <- 175
numForCompleteBP1 <- round( ( maxCompleteDOYBP1-minCompleteDOYBP1 ) * 0.9 )
minCompleteDOYBP3 <- 225
maxCompleteDOYBP3 <- 350
numForCompleteBP3 <- round( ( maxCompleteDOYBP3-minCompleteDOYBP3 ) * 0.9 )

#Number of days in a row that need to be within the CIs to get assigned synchronised:
#numForward <- 14 + 1
#numForward <- 7  + 0
numForwardSpring <- 10 + 0
numForwardFall   <- 16 + 0

for (j in 1:nSites){
  
  curSite <- which(e$site == siteList[j])
	e1 <- e[curSite,]

	e3Spring <- e1[ e1$dOY >= minCompleteDOYBP1 & e1$dOY <= maxCompleteDOYBP1, ]
	
	completeYearsSpring <- NULL #empty out from previous run
  
  if ( !empty( e3Spring ) ) {  #in case e3Spring is empty - avoids error
	  completeSpring <- as.data.frame( table( e3Spring$year,is.na( e3Spring$temp ) ) )
	  incompleteYearsSpring <- as.numeric(as.character(completeSpring$Var1[completeSpring$Var2 == 'FALSE' & completeSpring$Freq <  numForCompleteBP1]))
	  completeYearsSpring <-   as.numeric(as.character(completeSpring$Var1[completeSpring$Var2 == 'FALSE' & completeSpring$Freq >= numForCompleteBP1]))
  }
  
	e3Fall <- e1[ e1$dOY >= minCompleteDOYBP3 & e1$dOY <= maxCompleteDOYBP3, ]
	
  completeYearsFall <- NULL
	
	if ( !empty( e3Fall ) ) {
	  completeFall <- as.data.frame( table( e3Fall$year,is.na( e3Fall$temp ) ) )
	  incompleteYearsFall <- as.numeric(as.character(completeFall$Var1[completeFall$Var2 == 'FALSE' & completeFall$Freq <  numForCompleteBP3]))
	  completeYearsFall <-   as.numeric(as.character(completeFall$Var1[completeFall$Var2 == 'FALSE' & completeFall$Freq >= numForCompleteBP3]))
	} 
  
	completeYearsSpringOrFall <- unique(c(completeYearsSpring,completeYearsFall))
	#quantiles <- data.frame(year=c(min(e1$year):max(e1$year)))
	#quantiles$lo <- NA
	#quantiles$hi <- NA 
	
	for (year in completeYearsSpringOrFall){ 

	  breaks$springOrFallBPComplete[ breaks$year == year & breaks$site == siteList[j] ] <- TRUE
	  print(c('BP 1 and 3',j,as.character(siteList[j]),year))
 
    lo <- breaks$quantileLo[breaks$year == year & breaks$site == siteList[j]] 
    hi <- breaks$quantileHi[breaks$year == year & breaks$site == siteList[j]] 
    

		runsSpring <- data.frame(array(NA,c(1,numForwardSpring)))
  	runsFall <- data.frame(array(NA,c(1,numForwardFall)))

    eYear <- e1[e1$year == year, ] 

		if(year %in% completeYearsSpring){
			
			#out <- segmented( lm(temp~dOY, data=e1[e1$year == year,]), seg.Z=~dOY, psi=list(dOY=c(100,200)))
			#tryCatch({out <- segmented( lm(temp~dOY, data=e1[e1$year == year,]), seg.Z=~dOY, psi=list(dOY=c(100,200)))},error=function(e){cat("ERROR :",conditionMessage(e), "\n")})

			# time forward until approx breakpoint ascending/descending water temp (in 'breaks')
			for (i in 1:(200)){ #-window*winFactor )){
				for (ii in 2:numForwardSpring ){
          
          if( (i+ii-2) %in% eYear$dOY ) {   # amke sure that the dOY is in the looping variable      
			  	  runsSpring[ i,ii ] <- 1*((eYear$movingMean[ eYear$dOY == (i+ii-2) ] >= lo) & (eYear$movingMean[ eYear$dOY == (i+ii-2) ] <= hi))
          } else
            runsSpring[ i,ii ] <- 1  
          
				}
				runsSpring[ i,1 ] <- prod( runsSpring[ i, 2:numForwardSpring ] )
			}
			breaks$springBP[ breaks$year == year & breaks$site == siteList[j] ] <- min(which(runsSpring[,1] == 1))
			breaks$springBPComplete[ breaks$year == year & breaks$site == siteList[j] ] <- TRUE
		} #completeYearsSpring loop
      
	 if(year %in% completeYearsFall){
			# Make this not arbitrary
			# time backward until approx breakpoint ascending/descending water temp (in 'breaks')
	   stopLoop <- max( c( minCompleteDOYBP3,min(eYear$dOY)+numForwardFall + 1 ) )  # to keep from going past lower limit if dOY
			for (i in  max(eYear$dOY):stopLoop){
				for (ii in 2:numForwardFall ){
          
				  if( (i-ii+2) %in% eYear$dOY ) { 
					  runsFall[ i,ii ] <- 1*((eYear$movingMean[ eYear$dOY == (i-ii+2) ] >= lo) & (eYear$movingMean[ eYear$dOY == (i-ii+2) ] <= hi))
				  } else
				    runsFall[ i,ii ] <- 1 
				}
				runsFall[ i,1 ] <- prod( runsFall[ i, 2:numForwardFall ] )
			}
			breaks$fallBP[ breaks$year == year & breaks$site == siteList[j] ] <- max(which(runsFall[,1] == 1))
			breaks$fallBPComplete[ breaks$year == year & breaks$site == siteList[j] ] <- TRUE
		}	#completeYearsFall loop
			
	} #completeYearsSpringOrFall loop
} #site loop

# add in means for missing BP. use the xBPWMean below to constrain the date range for estimating BP 2 

#Add watershed or some spatial grouping to mean BPs.
# adapt to do things by BP & HUC 8

# Get the mean BPs for each site:
meanBP <- ddply( breaks, .(site), summarise, meanSpringBP = mean(springBP,na.rm=T), meanFallBP = mean(fallBP,na.rm=T) )
breaks$meanSpringBP <- NULL; breaks$meanFallBP <- NULL # in case we've merged before

breaks <- merge( x=breaks, y= meanBP, all.x=T, by=c('site') )

# Get the mean BPs for each HUC:
#  Currently means over all site/year combos. Other option would be to average all site means.
meanBPHUC8 <- ddply( breaks, .(HUC8), summarise, meanSpringBPHUC8 = mean(springBP,na.rm=T), meanFallBPHUC8 = mean(fallBP,na.rm=T) )
breaks$meanSpringBPHUC8 <- NULL; breaks$meanFallBPHUC8 <- NULL # in case we've merged before

breaks <- merge( x=breaks, y= meanBPHUC8, all.x=T, by=c('HUC8') )

#Define the averages of the calculated breakpoints.
meanSpringBP <- mean(breaks$meanSpringBP[is.finite(breaks$meanSpringBP)])
meanFallBP   <- mean(breaks$meanFallBP[is.finite(breaks$meanFallBP)])

#If the breakpoint doesn't exist, replace it with the HUC8 mean. (Same if it's been calculated as NaN.)
breaks$springBPWMean <- ifelse( is.na(breaks$springBP), breaks$meanSpringBPHUC8, breaks$springBP )
breaks$springBPWMean <- ifelse( !is.finite(breaks$springBPWMean), breaks$meanSpringBPHUC8, breaks$springBPWMean )
breaks$fallBPWMean   <- ifelse( is.na(breaks$fallBP), breaks$meanFallBPHUC8, breaks$fallBP )
breaks$fallBPWMean   <- ifelse( !is.finite(breaks$fallBPWMean), breaks$meanFallBPHUC8, breaks$fallBPWMean )

#If the breakpoint still doesn't exist, replace it with the overall mean. (Same if it's been calculated as NaN.)
breaks$springBPWMean <- ifelse( is.na(breaks$springBPWMean), meanSpringBP, breaks$springBPWMean )
breaks$springBPWMean <- ifelse( !is.finite(breaks$springBPWMean), meanSpringBP, breaks$springBPWMean )
breaks$fallBPWMean <- ifelse( is.na(breaks$fallBPWMean), meanFallBP, breaks$fallBPWMean )
breaks$fallBPWMean <- ifelse( !is.finite(breaks$fallBPWMean), meanFallBP, breaks$fallBPWMean )

# can have empty breakPoints if don't calculate a mean across years for each site in the ddply above
# i.e. only data for one year in a site or no years have data before the spring cutoff or after the fall cutoff
# in this case, just fill in the cutoffs, so the summer breakPoint can be calculated
breaks$springBPWMean[is.na(breaks$springBPWMean )] <- minCompleteDOYBP1
breaks$fallBPWMean  [is.na(breaks$fallBPWMean   )] <- maxCompleteDOYBP3
```

```{r merge breaks}
#----------------------------------------------------------------
#Merge the breakpoints and quantiles with the existing dataframe:
#----------------------------------------------------------------
#et <- merge( x=e, y=breaks, by.x=c('year','site'),by.y=c('year','site'),all.x=T)
et <- merge( x=e, y=breaks, by=c('year','site'),all.x=T)

# need to sort for plotting pred
#segDF <- segDF[order(segDF$site,segDF$year,segDF$dOY),]

```


```{r save e}
 # save(e,                       file=paste0(dataOutDir, sourceList, '/', 'e_' , sourceList, '.RData'))
  save(et, breaks, file=paste0(dataOutDir, '/', 'ets', '.RData'))

```

