Overall approach: 
1) Define bps and slopes by segment for each site/year combo in temperatureSegmentedBreakPointAnalysis.Rmd. 
2) Model slopes for each segment (2=sp-sum, 3=sum-autumn) as a function of airTemp and fixed covariates. This gtempDataSyncS predicted water temp as a function of airTemp and covariates, but does not identify bps.
3) Model bps as a fucntion of covariates including swe for bp1.
4) Predict water temp as function of airTemp and covariates btempDataSyncween bps for each prediction site
5) Summarize data for slopes btw bps

Note: run temperatureSegmentedBreakPointAnalysis.Rmd before running this script

```{r load libs}
rm(list=ls())

library(ggplot2)
library(dplyr)
library(DataCombine) # for the slide function
library(nlme)

#stempDataSyncwd('/Users/Dan/Documents/Research/Stream_Climate_Change/temperatureProject/')
#stempDataSyncwd('C:/Users/dhocking/Documents/temperatureProject/')

#baseDir <- 'C:/KPONEIL/GitHub/projects/temperatureProject/'
#baseDir <- '/Users/Dan/Documents/Research/Stream_Climate_Change/temperatureProject/'
baseDir <- 'C:/Users/dhocking/Documents/temperatureProject/'

dataInDir <- paste0(baseDir, 'dataIn/')
dataOutDir <- paste0(baseDir, 'dataOut/')
dataLocalDir <- paste0(baseDir, 'localData')
graphsDir <- paste0(baseDir, 'graphs/')

source(paste0(baseDir, 'code/functions/temperatureModelingFunctions.R'))
source(paste0(baseDir, 'code/functions/dataIndexingFunctions.R'))

load(paste0(dataOutDir, 'springFallBreakpoints.RData'))

# r Choose data source

#Northeast
CTDEP  <- T
MAFW   <- T
MAUSGS <- T
MADEP  <- T 
NHFG   <- F
NHDES  <- F
USFS   <- F
VTFWS  <- F
MEDMR  <- F

#Montana
MTUSGSYellowstone <- F
MTUSGSGlacier <- F

sourceChoice <- list( CTDEP,   MAFW,   MAUSGS, MADEP,   NHFG,   NHDES,   MEDMR,   USFS,   VTFWS,    MTUSGSYellowstone,   MTUSGSGlacier)
sourceNames  <- c   ('CTDEP', 'MAFW', 'MAUSGS', 'MADEP', 'NHFG', 'NHDES', 'MEDMR', 'USFS', 'VTFWS',  'MTUSGSYellowstone', 'MTUSGSGlacier')

dataSource <- sourceNames[sourceChoice == T]

fields <- c("agency", "date", "AgencyID", "year", "site", "date", "dOY", "temp", "airTemp", "prcp", "srad", "dayl", "swe")

tempData <- readStreamTempData(timeSeries=TRUE, covariates=TRUE, dataSourceList=dataSource, fieldListTS=fields, fieldListCD='ALL', directory=dataInDir)
springFallBPs$site <- as.character(springFallBPs$site)

# Join with break points
tempDataBP <- left_join(tempData, springFallBPs, by=c('site', 'year'))
rm(tempData) # save some memory

# Clip to syncronized season
tempDataSync <- filter(tempDataBP, dOY >= finalSpringBP & dOY <= finalFallBP)

# Creat site and year factor variables
tempDataSync$fyear <- as.factor(tempDataSync$year)
tempDataSync$fsite <- as.factor(tempDataSync$site)
```

```{r lag airTemp & prcp}

# Order by group and date
tempDataSync <- tempDataSync[order(tempDataSync$site,tempDataSync$year,tempDataSync$dOY),]

# For checking the order of tempDataSync
tempDataSync$count <- 1:length(tempDataSync$year)

tempDataSync <- tempDataSync[order(tempDataSync$count),] # just to make sure tempDataSync is ordered for the slide function

# airTemp
tempDataSync <- slide(tempDataSync, Var = "airTemp", GroupVar = "site", slideBy = -1, NewVar='airTempLagged1')
tempDataSync <- slide(tempDataSync, Var = "airTemp", GroupVar = "site", slideBy = -2, NewVar='airTempLagged2')

# prcp
tempDataSync <- slide(tempDataSync, Var = "prcp", GroupVar = "site", slideBy = -1, NewVar='prcpLagged1')
tempDataSync <- slide(tempDataSync, Var = "prcp", GroupVar = "site", slideBy = -2, NewVar='prcpLagged2')
tempDataSync <- slide(tempDataSync, Var = "prcp", GroupVar = "site", slideBy = -3, NewVar='prcpLagged3')

```

```{r}
# Make dataframe with just variables for modeling and order before standardizing
tempDataSync <- tempDataSync[ , c("agency", "date", "AgencyID", "year", "fyear", "site", "fsite", "date", "finalSpringBP", "finalFallBP", "FEATUREID", "HUC4", "HUC8", "HUC12", "temp", "Latitude", "Longitude", "airTemp", "airTempLagged1", "airTempLagged2", "prcp", "prcpLagged1", "prcpLagged2", "prcpLagged3", "dOY", "Forest", "Herbacious", "Agriculture", "Developed", "TotDASqKM", "ReachElevationM", "ImpoundmentsAllSqKM", "HydrologicGroupAB", "SurficialCoarseC", "CONUSWetland", "ReachSlopePCNT", "srad", "dayl", "swe")] #  

summary(tempDataSync)
dim(tempDataSync)
tempDataSync <- na.omit(tempDataSync) ####### Change this so don't take out NA in stream temperature
dim(tempDataSync)
```

```{r separate validation data}
#Use validation?
validate = T
  
#If validating:
  # Choose fraction of total # of sites:
validateFrac <- 0.2

if(validate) {
  n.fit <- floor(length(unique(tempDataSync$site)) * (1 - validateFrac))

  set.seed(2346)
  site.fit <- sample(unique(tempDataSync$site), n.fit, replace = FALSE) # select sites to hold back for testing 
  tempDataSyncValid <- subset(tempDataSync, !site %in% site.fit) # data for validation
  tempDataSync <- subset(tempDataSync, site %in% site.fit)    # data for model fitting (calibration)
  } else {
    tempDataSyncValid <- NULL
  }

```

```{r Standardize}
# Standardize for Analysis

tempDataSyncS <- cbind(tempDataSync[ ,c(1:15)],
             apply(X = tempDataSync[ ,16:dim(tempDataSync)[2]], MARGIN=2,
                   FUN = function(x){(x-mean(x, na.rm=TRUE)) / sd(x, na.rm=TRUE)}))



tempDataSyncValidS <- cbind(tempDataSyncValid[ ,c(1:15)],
             apply(X = tempDataSyncValid[ ,16:dim(tempDataSyncValid)[2]], MARGIN=2,
                   FUN = function(x){(x-mean(x, na.rm=TRUE)) / sd(x, na.rm=TRUE)}))

tempDataSyncValidS$fyear <- as.factor(tempDataSyncValidS$year)
tempDataSyncValidS$fsite <- as.factor(tempDataSyncValidS$site)



```


```{r save tempDataSync for use in analysis}

save(tempDataSync, tempDataSyncS, tempDataSyncValid, tempDataSyncValidS, file=paste0(dataOutDir, 'tempDataSync.RData'))

```

Left out to save time:
----------------------------------------------------------------------------------------------------------------------------------------------------------------
# 5-day mean of prcp 
siteYearCombos <- unique(tempDataSync[,c('site','year')])

tempDataSync$prcp5Day <- NA

window <- 5
for (i in 1:nrow(siteYearCombos)){

  print(c(i,as.character(siteYearCombos$site[i]),siteYearCombos$year[i],i/nrow(siteYearCombos)))
  
  currSite <- which(tempDataSync$site == as.character(siteYearCombos$site[i]) & tempDataSync$year == siteYearCombos$year[i] )

  #Need this so sites with very short records don't crash the loop.
  if(length(currSite) >= window){currMean <-  rollapply(tempDataSync$prcp[currSite], width=window, fill=NA, mean, align = 'left')} else(currMean <- NA)
  
  tempDataSync$prcp5Day[currSite] <- currMean
}
----------------------------------------------------------------------------------------------------------------------------------------------------------------


