Overall approach: 
1) Define bps and slopes by segment for each site/year combo in temperatureSegmentedBreakPointAnalysis.Rmd. 
2) Model slopes for each segment (2=sp-sum, 3=sum-autumn) as a function of airTemp and fixed covariates. This gets predicted water temp as a function of airTemp and covariates, but does not identify bps.
3) Model bps as a fucntion of covariates including swe for bp1.
4) Predict water temp as function of airTemp and covariates between bps for each prediction site
5) Summarize data for slopes btw bps

```{r load libs}
rm(list=ls())

library(ggplot2)
library(relaimpo)
library(lme4)
library(DataCombine) # for the slide function
library(plyr)
library(reshape)
library(ggmap)
library(foreign)
library(maptools)

baseDir <- 'C:/KPONEIL/'

dataInDir <- paste0(baseDir, 'GitHub/projects/temperatureProject/dataIn/')
dataOutDir <- paste0(baseDir, 'GitHub/projects/temperatureProject/dataOut/')
graphsDir <- paste0(baseDir, 'GitHub/projects/temperatureProject/graphs/')

```

Which agencies do you want to pull data from?
```{r Choose data source}

CTDEP  <- T
MAFW   <- T
MAUSGS <- T
NHFG   <- T
NHDES  <- T
MEDMR  <- T
USFS   <- T
VTFWS  <- T
MTUSGSYellowstone <- F
MTUSGSGlacier <- F

sourceChoice <- list( CTDEP,   MAFW,   MAUSGS,   NHFG,   NHDES,   MEDMR,   USFS,   VTFWS,    MTUSGSYellowstone,   MTUSGSGlacier)
sourceNames  <- c   ('CTDEP', 'MAFW', 'MAUSGS', 'NHFG', 'NHDES', 'MEDMR', 'USFS', 'VTFWS',  'MTUSGSYellowstone', 'MTUSGSGlacier')
  
dataSource <- sourceNames[sourceChoice == T]

sourceList <- paste0(paste0(dataSource, collapse = '_'))#, '_', Sys.Date())

```


```{r global vars}
dpiIn <- 400
```

```{r load pre-run dataframes}

#load(paste0(basedir, 'data/temperature/fromKyle/CTday2010AfterR.RData'))
#load(paste0(basedir, 'data/temperature/fromKyle/UpstreamStatsCTAfterR.RData'))

```


```{r load data}
# run temperatureSegmentedBreakPointAnalysis.Rmd before running this script

#Load "et" for the agencies:
load(paste0(dataOutDir, sourceList,  '/et_', sourceList, '.RData'))

#These are duplicates that get removed from daymet dataframe before merging:
removed1 <- c('Latitude', 'Longitude')
removed2 <- c('AgencyID', 'agency', 'temp', 'airTemp', 'Latitude', 'Longitude')

#Load in covariate data to merge into slopes df [no day data]
for ( i in 1:length(dataSource)){

  #Load daymet climate data to be merged into et:
  load(paste0(dataInDir, dataSource[i], '/streamTempSitesObservedClimateData_', dataSource[i], '.RData')) 
  masterData <- masterData[, -which(names(masterData) %in% removed2)]
  if ( i == 1) {newDay <- masterData} else ( rbind(newDay, masterData) )
  
  #Load covariate data to be merged into slopes df [no day data]
  load(paste0(dataInDir, dataSource[i], '/covariateData_', dataSource[i], '.RData')) #Jan 2014 covariateData [fixed over time] 
  covariateData <- covariateData[,-which(names(covariateData) %in% removed1)]
  if ( i == 1) {covs <- covariateData} else (covs <- rbind(covs, covariateData))

}
masterData    <- newDay
covariateData <- covs


et <- merge(et, masterData, by = c('site', 'year', 'dOY'), all.x=T, sort = F )

et$flow <- NA 
et$tAirMin <- et$tmin; et$tAirMax <- et$tmax 


#Overwrite NaNs with NAs:
covariateData <- replace(covariateData, is.na(covariateData), NA)

#covariateData$site <- covariateData$Site
covariateData$site <- as.character(covariateData$site)

et <- merge(et, covariateData, by = 'site', all.x=T, sort = F )

# get bps out of et
bp <- unique(et[,c('site','year','springBP','summerBP','fallBP')]  )
bp <- bp[is.finite(bp$springBP) | is.finite(bp$summerBP) | is.finite(bp$fallBP),]
bp$site <- as.character(bp$site) #for merging
  
siteData <- merge( x = bp, y = covariateData, by = 'site', all.x=T )
siteData$site <- siteData$Site

# turn Inf to NA in bps
siteData[!is.finite(siteData$springBP),'springBP'] <- NA
siteData[!is.finite(siteData$summerBP),'summerBP'] <- NA
siteData[!is.finite(siteData$fallBP),'fallBP'] <- NA

# merge in count of days
obsBySiteYear <- ddply(et, .(site,year), summarize,count=length(!is.na(temp)))
siteData <- merge(x=siteData, y=obsBySiteYear, all.x=T)

```

```{r lag airTemp}
et <- et[order(et$count),] # just to make sure et is ordered for the slide function

et <- slide(et, Var = "airTemp", GroupVar = "site", slideBy = -1, NewVar='airTempLagged1')
et <- slide(et, Var = "airTemp", GroupVar = "site", slideBy = -2, NewVar='airTempLagged2')
```

```{r scale daymet vars}
#don't use "scale" because it creates a data type with attributes that make is difficult to use predict
et$daylS <- (et$dayl-mean(et$dayl,na.rm=T))/sd(et$dayl,na.rm=T)#scale(et$dayl)
et$sradS <- (et$srad-mean(et$srad,na.rm=T))/sd(et$srad,na.rm=T)#scale(et$srad)
et$sweS <-   (et$swe-mean(et$swe,na.rm=T))/sd(et$swe,na.rm=T)#scale(et$swe)
```


```{r check out data}
#pairs(~Latitude+Longitude+Forest+ Impervious+ Agriculture+ BasinElevationM+ ReachSlopePCNT+ TotDASqKM+ WetlandOrWater+ SurficialCoarseC,data=et)
#Latitude, Longitude, Forest, Impervious, Agriculture, BasinElevationM, ReachSlopePCT, TotDASqKM, WetlandOrWater, SurficialCoarseC


#Makes barcode looking plot of data records:
#-------------------------------------------
gTile <- 
ggplot(siteData,aes(site,year,z=any(c(!is.na(springBP),!is.na(summerBP),!is.na(fallBP)))))+
  scale_x_discrete('Site')+
  scale_y_continuous('Year')+
  theme_bw(base_size=20) + 
    theme(axis.text.x = element_blank())+
  geom_tile()

ggsave( file=paste0(graphsDir, sourceList, '/gTile', sourceList, '.png'), plot=gTile, dpi=dpiIn , width=8,height=5, units='in' )


#Colors by number of observations?
#---------------------------------
gTileHeat <- 
ggplot(siteData,aes(site,year,z=count))+
  geom_tile(aes(fill=count))+
  scale_x_discrete('Site')+
  scale_y_continuous('Year')+ 
  theme(axis.text.x = element_blank())

ggsave( file=paste0(graphsDir, sourceList,'/gTileHeat', sourceList, '.png'), plot=gTileHeat, dpi=dpiIn , width=8,height=5, units='in' )
```

```{r regression temp~airTemp+...}
#Define covariates used in model. This gets used later in the prediction stage as well.
fixCov <- c("Latitude","Longitude","Forest", "Agriculture","BasinElevationM","ReachSlopePCNT", "TotDASqKM", "WetlandOrWater","SurficialCoarseC", "ImpoundmentsOpenSqKM", "WetlandsOpenSqKM")

et2 <- et[et$segment %in% 2,]
et3 <- et[et$segment %in% 3,]

cor(et3[,fixCov])

# Segment 2:
#===========
m0S2 <- lm(temp~airTemp+airTempLagged1+airTempLagged2,
         data=et2)

m1S2 <- lm(temp~airTemp+airTempLagged1+airTempLagged2+
              #segment+
              Latitude+Longitude+
              Forest+ Agriculture+ 
              BasinElevationM+ ReachSlopePCNT+ TotDASqKM+ 
              WetlandOrWater+ SurficialCoarseC +
              daylS + sradS + sweS,
         data=et2)

m2S2 <- lm(temp~(airTemp+airTempLagged1+airTempLagged2+
              #segment+
              Latitude+Longitude+
              Forest+ Agriculture+ 
              BasinElevationM+ ReachSlopePCNT+ TotDASqKM+ 
              WetlandOrWater+ SurficialCoarseC+
              daylS + sradS + sweS)^2,
         data=et2)

m3S2 <- lm(temp~(airTemp+airTempLagged1+airTempLagged2+
              #segment+
              Latitude+Longitude+
              Forest+ Agriculture+ 
              BasinElevationM+ ReachSlopePCNT+ TotDASqKM+ 
              WetlandOrWater+ SurficialCoarseC+ 
              ImpoundmentsOpenSqKM + WetlandsOpenSqKM +
              daylS + sradS + sweS)^2,
         data=et2)

AIC(m0S2,m1S2,m2S2,m3S2)


# Segment 3:
#===========
m0S3 <- lm(temp~airTemp+airTempLagged1+airTempLagged2,
         data=et3)

m1S3 <- lm(temp~airTemp+airTempLagged1+
              #segment+
              Latitude+Longitude+
              Forest+ Agriculture+ 
              BasinElevationM+ ReachSlopePCNT+ TotDASqKM+ 
              WetlandOrWater+ SurficialCoarseC+
              daylS + sradS + sweS,
         data=et3)

m2S3 <- lm(temp~(airTemp+airTempLagged1+airTempLagged2+
              #segment+
              Latitude+Longitude+
              Forest+ Agriculture+ 
              BasinElevationM+ ReachSlopePCNT+ TotDASqKM+ 
              WetlandOrWater+ SurficialCoarseC +
              daylS + sradS + sweS)^2,
         data=et3)


m3S3 <- lm(temp~(airTemp+airTempLagged1+airTempLagged2+
              #segment+
              Latitude+Longitude+
              Forest+ Agriculture+ 
              BasinElevationM+ ReachSlopePCNT+ TotDASqKM+ 
              WetlandOrWater+ SurficialCoarseC+ 
              ImpoundmentsOpenSqKM + WetlandsOpenSqKM +
              daylS + sradS + sweS)^2,
         data=et3)


AIC(m0S3,m1S3,m2S3,m3S3)
#r3 <- boot.relimp(m2S3)

```


```{r Interactions specified}
lm(temp~airTemp + airTempLagged1 + airTempLagged2 + Latitude + Longitude + BasinElevationM + ReachSlopePCNT + 
     TotDASqKM + SurficialCoarseC + Forest + Agriculture + WetlandOrWater + ImpoundmentsOpenSqKM + WetlandsOpenSqKM +
     daylS + sradS + sweS + 
     
     airTemp*ReachSlopePCNT + airTemp*TotDASqKM + airTemp*WetlandOrWater + 
     airTemp*ImpoundmentsOpenSqKM + airTemp*WetlandsOpenSqKM + airTemp*sweS + 
     
     airTempLagged1 + airTempLagged1*ReachSlopePCNT + airTempLagged1*TotDASqKM + airTempLagged1*WetlandOrWater + 
     airTempLagged1*ImpoundmentsOpenSqKM + airTempLagged1*WetlandsOpenSqKM + airTempLagged1*sweS + 
     
     airTempLagged2 + airTempLagged2*ReachSlopePCNT + airTempLagged2*TotDASqKM + airTempLagged2*WetlandOrWater + 
     airTempLagged2*ImpoundmentsOpenSqKM + airTempLagged2*WetlandsOpenSqKM + airTempLagged2*sweS + 
   
     ReachSlopePCNT*ImpoundmentsOpenSqKM + ReachSlopePCNT*WetlandsOpenSqKM + ReachSlopePCNT*daylS + ReachSlopePCNT*sradS +
   
     TotDASqKM*ImpoundmentsOpenSqKM + TotDASqKM*WetlandsOpenSqKM + TotDASqKM*daylS + TotDASqKM*sradS +
   
     SurficialCoarseC*Forest +   
     
     Forest*daylS + Forest*sradS + Forest*sweS +
     
     WetlandOrWater*daylS + WetlandOrWater*sradS + 

     ImpoundmentsOpenSqKM*daylS + ImpoundmentsOpenSqKM*sradS + 
     WetlandsOpenSqKM*daylS + WetlandsOpenSqKM*sradS +
     
     daylS*sweS + 
     sradS*sweS,
         data=et3)


```



```{r get predicted values}
et2$pred <- predict(m3S2, newdata=et2)
et2$pred0 <- predict(m0S2, newdata=et2)

et3$pred <- predict(m3S3, newdata=et3)
et3$pred0 <- predict(m0S3, newdata=et3)


```

```{r graphs of predicted/obs etc}
#all data
gPredObs <- 
  ggplot(et3,aes(pred,temp))+
  geom_point(size=0.5) +
    scale_x_continuous(expression(paste("Predicted water temperature (",degree, "C)", sep = "")))+
    scale_y_continuous(expression(paste("Observed water temperature (",degree, "C)", sep = "")))+
  theme_bw(base_size=20) +
  geom_abline(intercept=0,slope=1,color='white')

ggsave( file=paste0(graphsDir, sourceList,'/predObs', sourceList, '.png'), plot=gPredObs, dpi=dpiIn , width=8,height=5, units='in' )

```


Models for breakpoints
some early dates for spring and some late days for fall - need to see what's up with those site/years

#```{r get index for swe}
## surrogate for snow model. Get last day in spring where swe is x and first day in fall where swe is x
#ggplot(SWERecord[SWERecord$year %in% c(2008,2007,2009),],aes(dOY,swe))+
#  geom_line(aes(color=site))+
#  facet_wrap(~year)
#
## decided to use year as RE in lmer for now. will need to add in more mechanistic model in the future
#
#```

```{r regression bp1~+...}

# try RE model with year as RE

m5bp1 <- lmer(springBP~
              Latitude+Longitude+
              Forest+ Agriculture+ 
              BasinElevationM+ ReachSlopePCNT+ TotDASqKM+ 
              WetlandOrWater+ SurficialCoarseC +
              (1|year),
         data=siteData)
anova(m5bp1)
              #  BasinElevationM*Latitude+

m6bp1 <- lmer(springBP~
              (Latitude+Longitude+
              Forest+ Agriculture+ 
              BasinElevationM+ ReachSlopePCNT+ TotDASqKM+ 
              WetlandOrWater+ SurficialCoarseC)^2 +
              (1|year),
         data=siteData)

m7bp1 <- lmer(springBP~
              Latitude+Longitude+
              Forest+ Agriculture+ 
              BasinElevationM+ ReachSlopePCNT+ TotDASqKM+ 
              WetlandOrWater+ SurficialCoarseC +
              ImpoundmentsOpenSqKM + WetlandsOpenSqKM +
              (1|year),
         data=siteData)
anova(m7bp1)

AIC(m5bp1,m6bp1,m7bp1)

# not sure why need this [allow.new.levels=T] but throws an error otherwise
#may be because year is in the df
siteData$bp1Pred <- predict(m7bp1,newdata=siteData,allow.new.levels=T)
siteData$bp1PredAvgYear <- predict(m7bp1,newdata=siteData,REform=NA)

```

```{r regression bp2~+...}


# try RE model with year as RE

m5bp2 <- lmer(summerBP~
              Latitude+Longitude+
              Forest+ Agriculture+ 
              BasinElevationM+ ReachSlopePCNT+ TotDASqKM+ 
              WetlandOrWater+ SurficialCoarseC +
              (1|year),
         data=siteData)
anova(m5bp2)

m6bp2 <- lmer(summerBP~
              (Latitude+Longitude+
              Forest+ Agriculture+ 
              BasinElevationM+ ReachSlopePCNT+ TotDASqKM+ 
              WetlandOrWater+ SurficialCoarseC +
              (1|year))^2,
         data=siteData)

m7bp2 <- lmer(summerBP~
              (Latitude+Longitude+
              Forest+ Agriculture+ 
              BasinElevationM+ ReachSlopePCNT+ TotDASqKM+ 
              WetlandOrWater+ SurficialCoarseC +
              TNC_DamCount + ImpoundmentsOpenSqKM + WetlandsOpenSqKM +
              (1|year))^2,
         data=siteData)
anova(m7bp2)


AIC(m5bp2,m6bp2,m7bp2)

# not sure why need this [allow.new.levels=T] but throws an error otherwise
siteData$bp2Pred <- predict(m7bp2,newdata=siteData,allow.new.levels=T)
siteData$bp2PredAvgYear <- predict(m7bp2,newdata=siteData,REform=NA)
```

```{r regression bp3~+...}

# try RE model with year as RE

m5bp3 <- lmer(fallBP~
              Latitude+Longitude+
              Forest+ Agriculture+ 
              BasinElevationM+ ReachSlopePCNT+ TotDASqKM+ 
              WetlandOrWater+ SurficialCoarseC +
              (1|year),
         data=siteData)
anova(m5bp3)

m6bp3 <- lmer(fallBP~
              (Latitude+Longitude+
              Forest+ Agriculture+ 
              BasinElevationM+ ReachSlopePCNT+ TotDASqKM+ 
              WetlandOrWater+ SurficialCoarseC)^2 +
              (1|year),
         data=siteData)

m7bp3 <- lmer(fallBP~
              (Latitude+Longitude+
              Forest+ Agriculture+ 
              BasinElevationM+ ReachSlopePCNT+ TotDASqKM+ 
              WetlandOrWater+ SurficialCoarseC +
              TNC_DamCount + ImpoundmentsOpenSqKM + WetlandsOpenSqKM +
              (1|year))^2,
         data=siteData)
anova(m7bp3)



AIC(m5bp3,m6bp3,m7bp3)

# not sure why need this [allow.new.levels=T] but throws an error otherwise
siteData$bp3Pred <- predict(m7bp3,newdata=siteData,allow.new.levels=T)
siteData$bp3PredAvgYear <- predict(m7bp3,newdata=siteData,REform=NA)
```

```{r pred bp graph}
gObsPredBP1 <- 
ggplot(siteData[siteData$springBP>25,], aes(bp1Pred,springBP))+
  geom_point()+
  geom_abline(intercept=0,slope=1)+
  scale_x_continuous("Predicted spring breakpoint")+
  scale_y_continuous("Observed spring breakpoint")+
  theme_bw(base_size=20)

ggsave( file=paste0(graphsDir, sourceList,'/gObsPredBP1_', sourceList, '.png'), plot=gObsPredBP1, dpi=dpiIn , width=8,height=5, units='in' )

graphsDir, sourceList,'/gObsPredBP3_', sourceList, '.png'

#need to look into 2008, low observed values
gObsPredBP2 <- 
ggplot(siteData[siteData$summerBP>180&siteData$summerBP<240&siteData$year!=2008,], aes(bp2Pred,summerBP))+
  geom_point()+
  geom_abline(intercept=0,slope=1)+
  scale_x_continuous("Predicted summer breakpoint")+
  scale_y_continuous("Observed summer breakpoint")+
  theme_bw(base_size=20) 

ggsave( file=paste0(graphsDir, sourceList,'/gObsPredBP2_', sourceList, '.png'), plot=gObsPredBP2, dpi=dpiIn , width=8,height=5, units='in' )

gObsPredBP3 <- 
ggplot(siteData, aes(bp3Pred,fallBP))+
  geom_point()+
  geom_abline(intercept=0,slope=1)+
  scale_x_continuous("Predicted fall breakpoint")+
  scale_y_continuous("Observed fall breakpoint")+
  theme_bw(base_size=20)#+facet_wrap(~year) 

ggsave( file=paste0(graphsDir, sourceList,'/gObsPredBP3_', sourceList, '.png'), plot=gObsPredBP3, dpi=dpiIn , width=8,height=5, units='in' )

```


```{r bp stats}

siteData$bp1bp3 <- siteData$bp3Pred - siteData$bp1Pred
siteData$bp1bp3AvgYear <- siteData$bp3PredAvgYear - siteData$bp1PredAvgYear

save(siteData,file=paste0(dataOutDir, sourceList,'/siteDataWBPs_', sourceList, '.RData'))
```

Predict values for selected catchements
1) Predict breakpoints for UpstreamStats
2) Merge bps into Daymet files
3) Identify segements in Daymet files
3) Predict water temp for segments 2,3

```{r Define prediction region, year, covariates used, and catchments.}

#Pick the area you want to predict for. This is done by selection of daymet tiles:
# See Map Here: http://daymet.ornl.gov/sites/default/files/images/Tiles_on_LCC_projection_300dpi_labels.jpg

DaymetTiles <- c(11754, 11755, 11934, 11935)

Year <- 2010

#Load the 
load(paste0(dataInDir, 'NENY_CovariateData_2014-02-04.RData')

if( exists("fixCov")){ predictionCovs <- c('FEATUREID', fixCov)} else( print("Define prediction covariates."))

#predictionCovs <- c('FEATUREID', .............)

#Read in NHD catchments you want to predict for. This should fall within the boundaries of the daymet tiles above.
proj4.NHD  <- "+proj=longlat +ellps=GRS80 +datum=NAD83 +no_defs"
catchments <- readShapePoly ( "C:/KPONEIL/USGS/NHDPlusV2/Modified Versions/CT_Catchment.shp", proj4string=CRS(proj4.NHD))

features <- catchments$FEATUREID

```

```{r predicted values for select catchments}

#Here "UpstreamStatsCT" has become "predictionStats" and "CTday2010" has become "FullRecord".

#############################
# predict bps

predictionStats <- UpstreamStats[ ,names(UpstreamStats) %in% predictionCovs]

rm(UpstreamStats, LocalStats)

predictionStats$year <- Year # for bp predictions

#REform=NA uses no REs. Default is to use all REs (year in our case)
predictionStats$bp1 <- predict(m7bp1,newdata=predictionStats, allow.new.levels = T)#,REform=~(1|year))
predictionStats$bp2 <- predict(m7bp2,newdata=predictionStats)#,REform=~(1|year))
predictionStats$bp3 <- predict(m7bp3,newdata=predictionStats)#,REform=~(1|year))

predictionStats$bp1bp3 <- predictionStats$bp3 - predictionStats$bp1
##########################

for ( i in 1:length(DaymetTiles)){
  
  print(i)
  
  # Read in daily data for tile:
  #-----------------------------
  setwd("C:/KPONEIL/USGS/Stream Temperature/data/temperature/fromKyle/BP_Analysis/BP_Analysis/DaymetClimateData")
  
  load(paste0('NHD_DaymetTile_' , DaymetTiles[i], '_', Year, '.RData'))
  
  predictionStatsTile <- predictionStats[predictionStats$FEATUREID %in% FullRecord$FEATUREID, ]

  if(exists('features')){predictionStatsTile <- predictionStatsTile[predictionStatsTile$FEATUREID %in% features, ]}
  if(exists('features')){FullRecord <- FullRecord[FullRecord$FEATUREID %in% features,]}
  
  FullRecord$site <- FullRecord$FEATUREID

  #length(which(is.na(FullRecord$airTemp)))
  #length(FullRecord$airTemp)

  # Merge in bps and assign segments. Do here to make FullRecord smaller:
  #----------------------------------------------------------------------
  FullRecord <- merge( x=FullRecord, y=predictionStatsTile, all.x=T, by = 'FEATUREID' )
  
  #limit FullRecord to between bp1 and bp3 and assign segments
  FullRecord <- FullRecord[FullRecord$dOY > FullRecord$bp1 & 
                         FullRecord$dOY < FullRecord$bp3,]
  FullRecord$segment <- ifelse( FullRecord$dOY > FullRecord$bp1 & 
                               FullRecord$dOY <= FullRecord$bp2, 2,
                       ifelse( FullRecord$dOY > FullRecord$bp2 &
                               FullRecord$dOY <= FullRecord$bp3, 3, NA))
  
  # get rid of site/segment combos with all low # of obs
  counts <- ddply( FullRecord, .(site,segment), summarize, count=length(na.omit(airTemp)))
  FullRecord <- merge( x=FullRecord, y=counts, all.x=T )
  FullRecord <- FullRecord[ FullRecord$count > 3, ]
  ############
  
  #scale the daymet variables in the models based on original mean and sd
  FullRecord$daylS <- (FullRecord$dayl-mean(et$dayl,na.rm=T))/sd(et$dayl,na.rm=T)
  FullRecord$sradS <- (FullRecord$srad-mean(et$srad,na.rm=T))/sd(et$srad,na.rm=T)
  FullRecord$sweS <-   (FullRecord$swe-mean(et$swe,na.rm=T))/sd(et$swe,na.rm=T)
  
  #############################
  # lag air temp
  FullRecord <- FullRecord[order(FullRecord$FEATUREID,FullRecord$dOY),] # just to make sure FullRecord is ordered for the slide function
  
  # these are slow for big datasets...
  FullRecord <- slide(FullRecord, Var = "airTemp", GroupVar = "site", slideBy = -1, NewVar='airTempLagged1')
  FullRecord <- slide(FullRecord, Var = "airTemp", GroupVar = "site", slideBy = -2, NewVar='airTempLagged2')
  
  #############################
  # predict daily water temp
  
  FullRecord$predTemp <- ifelse( FullRecord$segment == 2, 
                                 predict(m3S2,FullRecord),
                        ifelse( FullRecord$segment == 3,        
                                 predict(m3S3,FullRecord), NA ) )

  # Estimate slopes of air/water for each site:
  #--------------------------------------------
  # Can't do this directly in the big regressions because other daily covariates besides airTemp are included (swe, dayl, srad)

  slopes <- ddply( FullRecord, .(site,segment), summarize, 
                       slope=coef(lm(predTemp ~ airTemp))[2])
  
  slopesMelt <- melt(slopes,id.vars=c('site','segment'))
  slopesCast <- cast(slopesMelt,site~segment)
  names(slopesCast) <- c('site','slopeSeg2','slopeSeg3')
  FullRecord <- merge( x=FullRecord, y=slopes, all.x=T, by = c('site','segment') )
  
  predictionStatsTile$site <- predictionStatsTile$FEATUREID
  predictionStatsTile <- merge( x=predictionStatsTile, y=slopesCast, all.x=T, by = c('site') )
  #ggplot(slopes, aes(slope)) + geom_histogram()+facet_wrap(~segment)
  
  if ( i == 1 ) {JoinFullRecord <- FullRecord} else(rbind(JoinFullRecord, FullRecord))
  
  if ( i == 1 ) {JoinPredictionStatsTile <- predictionStatsTile} else(JoinPredictionStatsTile <- rbind(JoinPredictionStatsTile, predictionStatsTile))
  
  
  TempPreds <- predictionStatsTile[,c('FEATUREID', 'bp1', 'bp2', 'bp3', 'bp1bp3', 'slopeSeg2', 'slopeSeg3')]
  
  if ( i == 1 ) {Predictions <- TempPreds}  else (Predictions <- rbind(Predictions, TempPreds))
  
}


paste0(dataOutDir, sourceList,'/siteDataWBPs_', sourceList, '.RData')

#Save Prediction Files:
#----------------------
save(JoinFullRecord, file=paste0(dataOutDir, sourceList, '/DaymetClimateDataNHD_DaymetTile' , DaymetTiles[i], '_', Year, 'AfterR_', sourceList, '.RData'))

save(JoinPredictionStatsTile, file=paste0(dataOutDir, sourceList, '/UpstreamStatsTile' , DaymetTiles[i], '_', Year, 'AfterR_', sourceList, '.RData'))


#Write out prediction files for ArcGIS:
#--------------------------------------
names(Predictions) <- c('FEATUREID', 'SpringBP', 'SummerBP', 'FallBP', 'SyncTemps', 'RiseSlope', 'FallSlope')
write.csv(Predictions, file = paste0(dataOutDir, sourceList, '/BP_Predictions', Year, '_', sourceList,'.csv'), row.names = F )

write.dbf(Predictions, file = paste0(dataOutDir, sourceList, '/BP_Predictions', Year, '_', sourceList,'.dbf'))

```


#```{r Explore prediction issues}

pred <- JoinPredictionStatsTile

MissSP <- pred[which(pred$bp1 <=0 | is.na(pred$bp1)),]

missing <- pred$FEATUREID[which(is.na(pred$PercentImpoundedOpen))]





load("C:/KPONEIL/USGS/GIS/Covariate Stats/NENY_CovariateData_2014-02-04.RData")

test <- which(is.na(UpstreamStats))
  
length(which(is.infinite(covariateData$PercentImpoundedOpen)))
length(which(is.infinite(PredicitonStats$ImpoundmentsOpenSqKM)))
covariateData$ImpoundmentsOpenSqKM


unique(JoinFullRecord$site[which(is.na(JoinFullRecord$airTemp))])


```
#
# A bunch of extra plots. Skip these for now.
#




#CHUNK 10

#```{r this section goes into chunk 10}

#Pick some sites here to look at. ( I think...)
#==============================================
ggplot(et3[et3$site %in% "USFS_2432688",],aes(airTemp,pred))+
  geom_point(aes(color=site)) +
  stat_smooth(method='lm')+
  facet_wrap(~year)
#siteIndex <- unique(et$site)

ggplot(et2[et2$site %in% "MAFW_MARoar55",],aes(pred,temp))+
  geom_point(aes(color=factor(year))) +
  geom_abline(intercept=0,slope=1,color='red')+
  facet_wrap(~year)

ggplot(et3[et3$site %in% "USFS_2432688",],aes(pred,temp))+
  geom_point(aes(color=factor(year))) +
  geom_abline(intercept=0,slope=1,color='red')+
  facet_wrap(~year)

gpredDOY <- 
ggplot(et3[et3$site %in% "MAFW_MARoar55",],aes(dOY,temp))+
  geom_line(aes(color=factor(year))) +
  geom_line(aes(dOY,pred)) +
  #geom_line(aes(dOY,airTemp),color='blue') +
  #geom_point(aes(dOY,airTemp),color='blue') +
  #geom_line(aes(dOY,pred0),color='orange') +
  #geom_abline(intercept=0,slope=1,color='red')+
  scale_x_continuous('Day of the year')+
  scale_y_continuous('Water temperature')+
  theme_bw(base_size=20) +
  theme(legend.position="none")+
  facet_wrap(~year)

ggsave( file=paste(basedir, 'data/temperature/fromKyle/gpredDOY_', IndexSource, '.png',sep=''), plot=gpredDOY, dpi=dpiIn , width=8,height=5, units='in' )


ggplot(et3[et3$site %in% "2432688",],aes(dOY,temp))+
  geom_line(aes(color=factor(year))) +
  geom_line(aes(dOY,airTemp)) +
  facet_wrap(~year)

```

















#```{r bp raw data graphs for talk}
dat <- et[ et$year == 2008 & et$site %in% '235' ,]

#hist(siteData$springBP)
#hist(siteData$summerBP)
#hist(siteData$fallBP)

#Histogram of breakpoints (dOY):
#-------------------------------
gBP <- 
ggplot(siteData, aes(springBP))+
  geom_histogram(fill='darkblue')+
  geom_histogram(aes(summerBP), fill='darkred')+
  geom_histogram(aes(fallBP), fill='darkgreen')+
  scale_x_continuous('Day of year',lim=c(50,365))+
  scale_y_continuous('Frequency')+
  theme_bw(base_size=20)

ggsave( file=paste(basedir, 'data/temperature/fromKyle/gBP_', IndexSource ,'.png',sep=''), plot=gBP, dpi=dpiIn , width=8,height=5, units='in' )


#   :
#-------------------------------
gMohsNoSmooth <- 
  ggplot( dat, aes(airTemp,temp))+
  geom_point() +
#  geom_smooth( method = 'nls', formula = y ~ m + ((a-m)/(1+exp(g*(b-x)))), se = F, start = list(a = 40,m = 0, g=1/15, b=10), size=1.25, colour='black') +
  theme_bw(base_size=20) +
  scale_x_continuous(expression(paste("Air temperature (",degree, "C)", sep = "")))+ 
  scale_y_continuous(expression(paste("Water temperature (",degree, "C)", sep = ""))) 

ggsave( file=paste(basedir, 'data/temperature/fromKyle/mohsNoSmooth', IndexSource ,'.png',sep=''), plot=gMohsNoSmooth, dpi=dpiIn , width=8,height=5, units='in' )

gMohs <- 
  ggplot( dat, aes(airTemp,temp))+
  geom_point() +
  geom_smooth( method = 'nls', formula = y ~ m + ((a-m)/(1+exp(g*(b-x)))), se = F, start = list(a = 40,m = 0, g=1/15, b=10), size=1.25, colour='black') +
  theme_bw(base_size=20) +
  scale_x_continuous(expression(paste("Air temperature (",degree, "C)", sep = "")))+ 
  scale_y_continuous(expression(paste("Water temperature (",degree, "C)", sep = ""))) 

ggsave( file=paste(basedir, 'D:/GitHub/data/temperature/fromKyle/mohs.png', sep = ""), plot=gMohs, dpi=dpiIn , width=8,height=5, units='in' )

gDOY <- 
  ggplot( dat,   aes(dOY,temp))+
  geom_point() +
  geom_line()+
  geom_point(aes(dOY,airTemp), color='red')+
  geom_line(aes(dOY,airTemp), color='red')+
  theme_bw(base_size=20) +
  theme(legend.position="none")+
  scale_x_continuous('Day of year')+ 
  scale_y_continuous(expression(paste("Temperature (",degree, "C)", sep = "")))

ggsave( file='D:/GitHub/data/temperature/fromKyle/gDOY.png', plot=gDOY, dpi=dpiIn , width=8,height=5, units='in' )


gSeg <- 
  ggplot( et[ et$year == 2008 & et$site %in% '235' &
              et$segment %in% 2:3  ,], aes(airTemp,temp,color=factor(segment)))+
  geom_point() +
  geom_smooth( method = 'lm', size=1.2,se=F) +
    scale_color_manual(values = c("1" = "blue","2" = "red","3" = "darkgreen"))+
  geom_point(data=et[ et$year == 2008 & et$site %in% '235' & et$segment %in% 1  ,], aes(airTemp,temp))+
  theme_bw(base_size=20) +
  theme(legend.position="none")+
  scale_x_continuous(expression(paste("Air temperature (",degree, "C)", sep = "")))+ 
  scale_y_continuous(expression(paste("Water temperature (",degree, "C)", sep = "")))

ggsave( file='D:/GitHub/data/temperature/fromKyle/gSeg.png', plot=gSeg, dpi=dpiIn , width=8,height=5, units='in' )

gSeg1to1 <- 
  ggplot( et[ et$year == 2008 & et$site %in% '235' &
              et$segment %in% 2:3  ,], aes(airTemp,temp,color=factor(segment)))+
  geom_point() +
  geom_abline(intercept=0,slope=1,size=1.25)+
  geom_smooth( method = 'lm', size=1.2,se=F) +
    scale_color_manual(values = c("1" = "blue","2" = "red","3" = "darkgreen"))+
  geom_point(data=et[ et$year == 2008 & et$site %in% '235' & et$segment %in% 1  ,], aes(airTemp,temp))+
  theme_bw(base_size=20) +
  theme(legend.position="none")+
  scale_x_continuous(expression(paste("Air temperature (",degree, "C)", sep = "")))+ 
  scale_y_continuous(expression(paste("Water temperature (",degree, "C)", sep = "")))

ggsave( file='D:/GitHub/data/temperature/fromKyle/gSeg1to1.png', plot=gSeg1to1, dpi=dpiIn , width=8,height=5, units='in' )

gSeg1to1b <- 
  ggplot( et[ et$year == 2011 & et$site %in% '2439192' &
                et$segment %in% 2:3  ,], aes(airTemp,temp,color=factor(segment)))+
  geom_point() +
  geom_abline(intercept=0,slope=1,size=1.25)+
  geom_smooth( method = 'lm', size=1.2,se=F) +
  scale_color_manual(values = c("1" = "blue","2" = "red","3" = "darkgreen"))+
  geom_point(data=et[ et$year == 2011 & et$site %in% '2439192' & et$segment %in% 1  ,], aes(airTemp,temp))+
  theme_bw(base_size=20) +
  theme(legend.position="none")+
  scale_x_continuous(expression(paste("Air temperature (",degree, "C)", sep = "")))+ 
  scale_y_continuous(expression(paste("Water temperature (",degree, "C)", sep = "")))

ggsave( file='D:/GitHub/data/temperature/fromKyle/gSeg1to1b.png', plot=gSeg1to1b, dpi=dpiIn , width=8,height=5, units='in' )

# 
gBP2 <- 
  ggplot( dat, aes((dOY),(movingMean))) +
  theme_bw(base_size=20) +
  geom_point() +
  geom_hline( aes(yintercept=quantileLo), colour='black') +
  geom_hline( aes(yintercept=quantileHi), colour='black') +
  geom_vline( aes(xintercept=as.numeric(springBP)),size=1.25) +
  geom_vline( aes(xintercept=as.numeric(fallBP)),size=1.25) +
  geom_vline( aes(xintercept=as.numeric(summerBP)),size=1.25) +
  scale_x_continuous('Day of year')  +
  scale_y_continuous('Temperature index', lim=c(-10,10))  

ggsave( file='D:/GitHub/data/temperature/fromKyle/gBP2.png', plot=gBP2, dpi=dpiIn , width=8,height=5, units='in',scale=2 )

gBP2Color <- 
  ggplot( dat, aes((dOY),(movingMean),color=factor(segment))) +
  theme_bw(base_size=30) +
  geom_point() +
  geom_hline( aes(yintercept=quantileLo), colour='black') +
  geom_hline( aes(yintercept=quantileHi), colour='black') +
  geom_vline( aes(xintercept=as.numeric(springBP)),size=1.25) +
  geom_vline( aes(xintercept=as.numeric(fallBP)),size=1.25) +
  geom_vline( aes(xintercept=as.numeric(summerBP)),size=1.25) +
    scale_color_manual(values = c("1" = "blue","2" = "red","3" = "darkgreen"))+
  theme(legend.position="none")+
  scale_x_continuous('Day of year')  +
  scale_y_continuous('Temperature index', lim=c(-10,10))  

ggsave( file='D:/GitHub/data/temperature/fromKyle/gBP2Color.png', plot=gBP2Color, dpi=dpiIn , width=8,height=5, units='in',scale=2 )


gBP2NoBP <- 
  ggplot( dat, aes((dOY),(movingMean))) +
  geom_point() +
  geom_hline( aes(yintercept=quantileLo), colour='black') +
  geom_hline( aes(yintercept=quantileHi), colour='black') +
#  geom_vline( aes(xintercept=as.numeric(springBP)),size=1.25) +
#  geom_vline( aes(xintercept=as.numeric(fallBP)),size=1.25) +
#  geom_vline( aes(xintercept=as.numeric(summerBP)),size=1.25) +
  scale_x_continuous('Day of year')  +
  scale_y_continuous('Temperature index', lim=c(-10,10)) +
  theme_bw(base_size=30) 

ggsave( file='D:/GitHub/data/temperature/fromKyle/gBP2NoBP.png', plot=gBP2NoBP, dpi=dpiIn , width=8,height=5, units='in',scale=2 )


```









___________________________________________________________________________________________________________________________________________________

                                                          BELOW HERE CAN BE RUN IN ARC
____________________________________________________________________________________________________________________________________________________









#```{r map data}
load(file=file=paste0(basedir, 'data/temperature/fromKyle/UpstreamStatsCTAfterR.RData'))

map <- qmap(c(lon=-72.8, lat=41.5), source="google", zoom=9)


# left/bottom/right/top
#map <- qmap(c(-73.4, 40.6,-71.5,41.6), source="google")

# slope segment 2
gSlopeSeg2 <- 
map +
    geom_point(aes(x = Longitude, y = Latitude, colour = slopeSeg2), data = UpstreamStatsCT[UpstreamStatsCT$slopeSeg2<1.1,],size=1.8) +
  #scale_colour_gradient2(low="red", high="blue")
  scale_colour_gradient(low = "blue", high = "red",limits=c(0.55, 1.)) 

ggsave( file=paste('D:/GitHub/data/temperature/fromKyle/gSlopeSeg2.png',sep=''), plot=gSlopeSeg2, dpi=dpiIn , width=6,height=5, units='in' )

#slope segment 3
gSlopeSeg3 <- 
map +
    geom_point(aes(x = Longitude, y = Latitude, colour = slopeSeg3), data = UpstreamStatsCT[UpstreamStatsCT$slopeSeg3<1.1,]) +
  scale_colour_gradient(low="red", high="blue",limits=c(0.55, 1.1))

ggsave( file=paste('D:/GitHub/data/temperature/fromKyle/gSlopeSeg3.png',sep=''), plot=gSlopeSeg3, dpi=dpiIn , width=6,height=5, units='in' )


# breakpoints
bp <- UpstreamStatsCT[UpstreamStatsCT$bp3<365&UpstreamStatsCT$bp1<365,]

gPredBP <- 
ggplot(bp, aes(bp1))+
  geom_histogram(fill='darkblue')+
  geom_histogram(aes(bp2), fill='darkred')+
  geom_histogram(aes(bp3), fill='darkgreen')+
  scale_x_continuous('Day of year')+#,lim=c(50,365))+
  scale_y_continuous('Frequency')+
  theme_bw(base_size=20)

ggsave( file=paste('D:/GitHub/data/temperature/fromKyle/gPredBP.png',sep=''), plot=gPredBP, dpi=dpiIn , width=8,height=5, units='in' )

#bp 1
gMapBP1 <- 
map +
    geom_point(aes(x = Longitude, y = Latitude, colour = bp1), data = bp[bp$bp1<95,]) +
  scale_colour_gradient(low="red", high="blue")

ggsave( file=paste('D:/GitHub/data/temperature/fromKyle/gMapBP1.png',sep=''), plot=gMapBP1, dpi=dpiIn , width=6,height=5, units='in' )

#bp 2
gMapBP2 <- 
map +
    geom_point(aes(x = Longitude, y = Latitude, colour = bp2), data = bp) +
  scale_colour_gradient(low="red", high="blue")

ggsave( file=paste('D:/GitHub/data/temperature/fromKyle/gMapBP2.png',sep=''), plot=gMapBP2, dpi=dpiIn , width=6,height=5, units='in' )

#bp 3
gMapBP3 <- 
map +
    geom_point(aes(x = Longitude, y = Latitude, colour = bp3), data = bp) +
  scale_colour_gradient(low="red", high="blue")

ggsave( file=paste('D:/GitHub/data/temperature/fromKyle/gMapBP3.png',sep=''), plot=gMapBP3, dpi=dpiIn , width=6,height=5, units='in' )

#bp1 bp3
gMapBP13 <- 
map +
    geom_point(aes(x = Longitude, y = Latitude, colour = bp1bp3), data = bp[bp$bp1<95,]) +
  scale_colour_gradient(low="red", high="blue")
  #theme(legend.position="none")

ggsave( file=paste('D:/GitHub/data/temperature/fromKyle/gMapBP13.png',sep=''), plot=gMapBP13, dpi=dpiIn , width=6,height=5, units='in' )



```

