# this file is for multiple sites, also works for a single site

# use of max vs min airTemp in addition to mean air temp

```{r load libraries}
rm(list=ls())

library(ggplot2)
library(GGally)
library(gridExtra)
library(reshape2)
library(mgcv)
library(nlme)
library(plyr)
library(segmented)
library(zoo)
library(ggmap)
library(pls)
library(MASS)
library(lme4)
library(DataCombine) # for the slide function

baseDir <- 'C:/KPONEIL/'

dataInDir <- paste0(baseDir, 'GitHub/projects/temperatureProject/dataIn/')
dataOutDir <- paste0(baseDir, 'GitHub/projects/temperatureProject/dataOut/')
graphsDir <- paste0(baseDir, 'GitHub/projects/temperatureProject/graphs/')

source(paste0(baseDir, 'GitHub/projects/temperatureProject/code/temperatureModelingFunctions.R'))
```

Which agencies do you want to pull data from?
```{r Choose data source}

CTDEP  <- T
MAFW   <- F
MAUSGS <- F
NHFG   <- F
NHDES  <- F
MEDMR  <- F
MTUSGS <- F

sourceChoice <- list( CTDEP,   MAFW,   MAUSGS,   NHFG,   NHDES,   MEDMR,   MTUSGS )
sourceNames  <- c   ('CTDEP', 'MAFW', 'MAUSGS', 'NHFG', 'NHDES', 'MEDMR', 'MTUSGS')
  
dataSource <- sourceNames[sourceChoice == T]

sourceList <- paste0(paste0(dataSource, collapse = '_'), '  ', Sys.Date())

setwd(paste0(baseDir, 'GitHub/projects/temperatureProject/'))

```

```{r rerun or load previous?}

reRunBP <- T

# Will need to specify date of creation for this to work:
if( !reRunBP ) {
  load(file=paste(baseDir, 'GitHub/projects/temperatureProject/e_', sourceList, '.RData',sep=''))
  load(file=paste(baseDir, 'GitHub/projects/temperatureProject/et_', sourceList, '.RData',sep=''))
}

```

```{r Make graphs?}
# settings variables
makeSpringFallGraphs <- T
indexAbnormalSummerGraphs <- F

#Creates folders for graphs if they don't exist:
#-----------------------------------------------
subGraphsDir <- paste0(graphsDir, sourceList)

if (file.exists(subGraphsDir)){
  setwd(file.path(subGraphsDir))
} else {
  dir.create(file.path(paste0(subGraphsDir)))
  dir.create(file.path(paste0(subGraphsDir, '/summerBP')))
  dir.create(file.path(paste0(subGraphsDir, '/springFallBP')))
  dir.create(file.path(paste0(subGraphsDir, '/problemSites')))
}

```

```{r Load data}
 if ( reRunBP ){

  setwd(dataInDir)

  for ( i in 1:length(dataSource)){
    load(paste0(dataSource[i], '/streamTempData_', dataSource[i], '.RData'))  #masterData
    if ( i == 1) {e <- masterData} else ( e <- rbind(e, masterData))
    rm(masterData)
  }

}
```

###############################################
# format for input file
# > str(e)
#'data.frame':   622507 obs. of  10 variables:
# $ site     : chr  "CTDEP_1" "CTDEP_1" "CTDEP_1" "CTDEP_1" ...
# $ year     : num  2008 2008 2008 2008 2008 ...
# $ dOY      : num  1 10 100 101 102 103 104 105 106 107 ...
# $ date     : Date, format: "2008-01-01" "2008-01-10" ...
# $ AgencyID : Factor w/ 1479 levels "1","1008","101",..: 1 1 1 1 1 1 1 1 1 1 ...
# $ agency   : chr  "CTDEP" "CTDEP" "CTDEP" "CTDEP" ...
# $ temp     : num  NA NA NA NA NA ...
# $ airTemp  : num  2.25 7.25 8 13.5 12.75 ...
# $ Latitude : num  41.2 41.2 41.2 41.2 41.2 ...
# $ Longitude: num  -73.3 -73.3 -73.3 -73.3 -73.3 ...
##############################################

# need to make sure that dOY is consecutive within a year_site combo, otherwise the slide function will line up non-consecutive dOYs
# If a df doesn't have consec dOY could use functions in zoo or do the indexing explicitly

```{r Data prep}

siteList <- unique(e$site)

#==============================================================================================
# This is the key metric for estimating the synchrony between air and water temp 
e$tempIndex <- (e$temp-e$airTemp)/(e$temp + 0.00000001) # add small # to avoid Inf
#==============================================================================================

# fill in missing dOY so sequence calcs (like moving mean, slope etc) 

# lag by group
e <- e[order(e$site,e$year,e$dOY),]

#For checking the order of e:
#----------------------------
e$count <- 1:length(e$year)

#Get the slope of the air/stream temperature relationship:
#---------------------------------------------------------
e$siteYear <- paste(e$site,e$year,sep='_')
e <- slide(e, Var = "temp", GroupVar = "siteYear", slideBy = +1)
e <- slide(e, Var = "airTemp", GroupVar = "siteYear", slideBy = +1)

e$rise <- e$temp1 - e$temp
e$run  <- e$airTemp1 - e$airTemp
e$slope <- e$rise / e$run
e$length <- sqrt(e$rise^2 + e$run^2)
e$angle <- atan(e$rise/e$run) * 180/pi # need to check radian units...

# for dOY changes
e$waterDelta <-    e$temp1 -    e$temp
e$airDelta   <- e$airTemp1 - e$airTemp

e <- e[order(e$count),]

```

```{r moving mean}
#===============================================================================
#Get moving mean and SD of temp index for each site and put into the data frame:
#===============================================================================
window <- 10 # frame sizefor moving mean, which is centered by default

nSites <- length(siteList)
siteYearCombos <- unique(e[,c('site','year')])
#siteYearCombos$site  <- factor(siteYearCombos$site)

e$movingMean <- NA
e$movingSD <- NA

for (i in 1:nrow(siteYearCombos)){

  print(c(i,as.character(siteYearCombos$site[i]),siteYearCombos$year[i],i/nrow(siteYearCombos)))
  
  currSite <- which(e$site == as.character(siteYearCombos$site[i]) & e$year == siteYearCombos$year[i] )

  if(length(currSite) >= window){currMean <-  rollapply(e$tempIndex[currSite], width=window, fill=NA, mean)} else(currMean <- NA)
  if(length(currSite) >= window){currSD <-    rollapply(e$tempIndex[currSite], width=window, fill=NA, sd)}   else(currSD <- NA)
  
  e$movingMean[currSite] <- currMean
  e$movingSD  [currSite] <- currSD
}

e$meanSDDiff <- e$movingSD - e$movingMean

# just to make sure the merge doens't screw up order
e <- e[order(e$count),]

# look at some raw data
#ggplot( e[e$site == e$site[2] ,], aes(dOY,temp) ) + geom_point() + geom_point(aes(dOY,airTemp),color='red') + facet_grid(site~year)
#table( e$year,e$site,is.na( e$temp ) )
```

```{r Create the breaks dataframe}

breaks <- data.frame(array(NA,c(nrow(siteYearCombos),7)))
names(breaks) <- c('site','year','springBP','summerBP','fallBP','quantileLo','quantileHi')
breaks$fallBPComplete <- FALSE
breaks$springBPComplete <- FALSE
breaks$springOrFallBPComplete <- FALSE
breaks$summerBPComplete <- FALSE

breaks$quantileLo <- NA
breaks$quantileHi <- NA

#Define breakpoint time period and range for tempIndex
beginningDayForCI <- 125
endingDayForCI <- 275
loCI <- 0.001
hiCI <- 0.999

for ( i in 1:nrow(siteYearCombos)){
  print(i)
  breaks$year[i] <- as.numeric(siteYearCombos$year[i])
  breaks$site[i] <- as.character(siteYearCombos$site[i]) #would like this to stay a factor, but it ends up as numeric in 'breaks'. not sure why
 
  tmp <- e[e$site == siteYearCombos$site[i] & e$year  %in% siteYearCombos$year[i] & e$dOY %in% beginningDayForCI:endingDayForCI,'tempIndex']
  if (any(!is.na(tmp))){
    TIQ <- quantile(tmp, probs=c(loCI,0.5,hiCI),na.rm=T)
    #  movingSDQ <- quantile(e[e$site == siteList[j] & e$year  %in% year & e$dOY %in% beginningDayForCI:endingDayForCI, 'movingSD'], probs=c(0.005,0.5,0.995),na.rm=T)
    
    breaks$quantileLo[i] <- TIQ[1]
    breaks$quantileHi[i] <- TIQ[3]
  }
}  

```

```{r Spring/fall bp}

#===================================================================================
#Use runs analysis of the movingMean to define spring and fall breakpoints:
#===================================================================================

minCompleteDOYBP1 <- 15
maxCompleteDOYBP1 <- 175
numForCompleteBP1 <- round( ( maxCompleteDOYBP1-minCompleteDOYBP1 ) * 0.9 )
minCompleteDOYBP3 <- 225
maxCompleteDOYBP3 <- 350
numForCompleteBP3 <- round( ( maxCompleteDOYBP3-minCompleteDOYBP3 ) * 0.9 )

#Number of days in a row that need to be within the CIs to get assigned synchronised:
#numForward <- 14 + 1 
#numForward <- 7  + 0 
numForwardSpring <- 10 + 0
numForwardFall   <- 16 + 0

for (j in 1:nSites){
  
  curSite <- which(e$site == siteList[j])
	e1 <- e[curSite,]

	e3Spring <- e1[ e1$dOY >= minCompleteDOYBP1 & e1$dOY <= maxCompleteDOYBP1, ]
	
	completeYearsSpring <- NULL #empty out from previous run
  
  if ( !empty( e3Spring ) ) {  #in case e3Spring is empty - avoids error
	  completeSpring <- as.data.frame( table( e3Spring$year,is.na( e3Spring$temp ) ) )
	  incompleteYearsSpring <- as.numeric(as.character(completeSpring$Var1[completeSpring$Var2 == 'FALSE' & completeSpring$Freq <  numForCompleteBP1]))
	  completeYearsSpring <-   as.numeric(as.character(completeSpring$Var1[completeSpring$Var2 == 'FALSE' & completeSpring$Freq >= numForCompleteBP1]))
  }
  
	e3Fall <- e1[ e1$dOY >= minCompleteDOYBP3 & e1$dOY <= maxCompleteDOYBP3, ]
	
  completeYearsFall <- NULL
	
	if ( !empty( e3Fall ) ) {
	  completeFall <- as.data.frame( table( e3Fall$year,is.na( e3Fall$temp ) ) )
	  incompleteYearsFall <- as.numeric(as.character(completeFall$Var1[completeFall$Var2 == 'FALSE' & completeFall$Freq <  numForCompleteBP3]))
	  completeYearsFall <-   as.numeric(as.character(completeFall$Var1[completeFall$Var2 == 'FALSE' & completeFall$Freq >= numForCompleteBP3]))
	} 
  
	completeYearsSpringOrFall <- unique(c(completeYearsSpring,completeYearsFall))
	#quantiles <- data.frame(year=c(min(e1$year):max(e1$year)))
	#quantiles$lo <- NA
	#quantiles$hi <- NA 
	
	for (year in completeYearsSpringOrFall){ 

	  breaks$springOrFallBPComplete[ breaks$year == year & breaks$site == siteList[j] ] <- TRUE
	  print(c('BP 1 and 3',j,as.character(siteList[j]),year))
 
    lo <- breaks$quantileLo[breaks$year == year & breaks$site == siteList[j]] 
    hi <- breaks$quantileHi[breaks$year == year & breaks$site == siteList[j]] 
    

		runsSpring <- data.frame(array(NA,c(1,numForwardSpring)))
  	runsFall <- data.frame(array(NA,c(1,numForwardFall)))

    eYear <- e1[e1$year == year, ] 

		if(year %in% completeYearsSpring){
			
			#out <- segmented( lm(temp~dOY, data=e1[e1$year == year,]), seg.Z=~dOY, psi=list(dOY=c(100,200)))
			#tryCatch({out <- segmented( lm(temp~dOY, data=e1[e1$year == year,]), seg.Z=~dOY, psi=list(dOY=c(100,200)))},error=function(e){cat("ERROR :",conditionMessage(e), "\n")})

			# time forward until approx breakpoint ascending/descending water temp (in 'breaks')
			for (i in 1:(200)){ #-window*winFactor )){
				for (ii in 2:numForwardSpring ){
          
          if( (i+ii-2) %in% eYear$dOY ) {   # amke sure that the dOY is in the looping variable      
			  	  runsSpring[ i,ii ] <- 1*((eYear$movingMean[ eYear$dOY == (i+ii-2) ] >= lo) & (eYear$movingMean[ eYear$dOY == (i+ii-2) ] <= hi))
          } else
            runsSpring[ i,ii ] <- 1  
          
				}
				runsSpring[ i,1 ] <- prod( runsSpring[ i, 2:numForwardSpring ] )
			}
			breaks$springBP[ breaks$year == year & breaks$site == siteList[j] ] <- min(which(runsSpring[,1] == 1))
			breaks$springBPComplete[ breaks$year == year & breaks$site == siteList[j] ] <- TRUE
		}
      
	 if(year %in% completeYearsFall){
			# Make this not arbitrary
			# time backward until approx breakpoint ascending/descending water temp (in 'breaks')
	   stopLoop <- max( c( minCompleteDOYBP3,min(eYear$dOY)+numForwardFall + 1 ) )  # to keep from going past lower limit if dOY
			for (i in  max(eYear$dOY):stopLoop){
				for (ii in 2:numForwardFall ){
          
				  if( (i-ii+2) %in% eYear$dOY ) { 
					  runsFall[ i,ii ] <- 1*((eYear$movingMean[ eYear$dOY == (i-ii+2) ] >= lo) & (eYear$movingMean[ eYear$dOY == (i-ii+2) ] <= hi))
				  } else
				    runsFall[ i,ii ] <- 1 
				}
				runsFall[ i,1 ] <- prod( runsFall[ i, 2:numForwardFall ] )
			}
			breaks$fallBP[ breaks$year == year & breaks$site == siteList[j] ] <- max(which(runsFall[,1] == 1))
			breaks$fallBPComplete[ breaks$year == year & breaks$site == siteList[j] ] <- TRUE
		}			
			
#		} #completeYears loop
	} #year loop
} #site loop

# add in means for missing BP. use the xBPWMean below to constrain the date range for estimating BP 2 

#Add watershed or some spatial grouping to mean BPs.
# adapt to do things by BP & HUC 8

meanBP <- ddply( breaks, .(site), summarise, meanSpringBP = mean(springBP,na.rm=T), meanFallBP = mean(fallBP,na.rm=T) )
breaks$meanSpringBP <- NULL; breaks$meanFallBP <- NULL # in case we've merged before

breaks <- merge( x=breaks, y= meanBP, all.x=T, by=c('site') )

meanSpringBP <- mean(breaks$meanSpringBP[is.finite(breaks$meanSpringBP)])
meanFallBP <- mean(breaks$meanFallBP[is.finite(breaks$meanFallBP)])

breaks$springBPWMean <- ifelse( is.na(breaks$springBP), breaks$meanSpringBP, breaks$springBP )
breaks$springBPWMean <- ifelse( !is.finite(breaks$springBPWMean), meanSpringBP, breaks$springBPWMean )

breaks$fallBPWMean <- ifelse( is.na(breaks$fallBP), breaks$meanFallBP, breaks$fallBP )
breaks$fallBPWMean <- ifelse( !is.finite(breaks$fallBPWMean), meanFallBP, breaks$fallBPWMean )

# can have empty breakPoints if don't calculate a mean across years for each site in the ddply above
# i.e. only data for one year in a site or no years have data before the spring cutoff or after the fall cutoff
# in this case, just fill in the cutoffs, so the summer breakPoint can be calculated
breaks$springBPWMean[is.na(breaks$springBPWMean )] <- minCompleteDOYBP1
breaks$fallBPWMean  [is.na(breaks$fallBPWMean   )] <- maxCompleteDOYBP3
```

```{r summer bp}

#====================================================================================================================
#Segmented regression of water temperature to get breakpoint between ascending and descending temps by year and site:
# summerBP or BP 2
# can get bad bp2 estimates when water temp increases very rapidly in the spring. as of 11/15/13, including the middle 90% of the datapts between bp1 and bp3 to trim the ends 
#====================================================================================================================
minCompleteDOYBP2 <- meanSpringBP
maxCompleteDOYBP2 <- meanFallBP
estimatedBP2 <- meanFallBP - meanSpringBP

numBeforeCompleteBP2 <- round( ( maxCompleteDOYBP2-minCompleteDOYBP2 ) * 0.35 )
numAfterCompleteBP2  <- round( ( maxCompleteDOYBP2-minCompleteDOYBP2 ) * 0.25 )

trimProp <- 0.05 # the proportion of obs btw bp1 and bp3 to trim for segmented

segDF <- data.frame(site=NA,year=NA,dOY=NA,temp=NA,pred=NA)

abnormalSiteYears <- data.frame(site=NA,year=NA)

for (j in 1:nSites){
  
  curSite <- which(e$site == siteList[j])

  e1 <- e[curSite,]
  
  for (year in unique(e1$year)){
    
    print(c('BP2',j,as.character(siteList[j]),year))

    e3 <- e1[ e1$dOY >= minCompleteDOYBP2 & e1$dOY <= maxCompleteDOYBP2 & e1$year == year, ]
    
    numBeforeBP2 <- length(which((e3$dOY[!is.na(e3$temp)] <= estimatedBP2) == TRUE))
    numAfterBP2  <- length(which((e3$dOY[!is.na(e3$temp)] >= estimatedBP2) == TRUE))
    
    #If the number of points before and after the estimated mean summer breakpoint exists, proceed to calculate site specific breakpoint.
    if(numBeforeBP2 > 0 & numAfterBP2 > 0){
      
      bp1 <- breaks$springBPWMean[breaks$site == siteList[j] & breaks$year == year ]  #Use similar method to do BPmean by HUC
      bp3 <- breaks$fallBPWMean[breaks$site == siteList[j] & breaks$year == year ]
      
      trim <- round((bp3-bp1) * trimProp)
      minDOY <- bp1 + trim
      maxDOY <- bp3 - trim
      
      if(exists('out')) rm(out) 
      dat <- e1[e1$year == year & e1$dOY >= minDOY & e1$dOY <= maxDOY,]
      tryCatch({out <- segmented( lm(temp~dOY, data=dat), 
                                  seg.Z=~dOY, psi=list(dOY=c(200)))},error=function(e){cat("ERROR :",conditionMessage(e), "\n")})      

      #Plot and save summer breakpoint if the criteria are met:
      #--------------------------------------------------------
      png(filename=paste0(subGraphsDir, '/summerBP/summerBP_',e1$site[1],'_',year,'.png'),width=1000, height=600, bg="white")
      plot(e1$dOY[e1$year %in% year], e1$temp[e1$year %in% year],ylim=c(-5,25),xlim=c(0,366))
       
      if(exists('out')){
        
        breaks$risingTemp  [breaks$year == year & breaks$site == siteList[j]] <- out$coeff[2]
        breaks$fallingTemp [breaks$year == year & breaks$site == siteList[j]] <- out$coeff[3]
        
        if( out$coeff[2] < 0 | out$coeff[3] > 0 & indexAbnormalSummerGraphs){
          mark <- data.frame(siteList[j], year)
          names(mark) <- c('site', 'year')
          abnormalSiteYears <- rbind(abnormalSiteYears, mark)       
        }
      
        if(numBeforeBP2 >= numBeforeCompleteBP2 & numAfterBP2 >= numAfterCompleteBP2 & out$coeff[2] > 0 & out$coeff[3] < 0 ){
        
          # For making seg regression graphs in ggplot. 
          segDF <- rbind( segDF, cbind(site=siteList[j],year=year,dOY=dat$dOY[!is.na(dat$temp)],temp=dat$temp[!is.na(dat$temp)],pred=out$fitted.values) )
        
          breaks$summerBP        [breaks$year == year & breaks$site == siteList[j]] <- summary(out)$psi[1,2]
          breaks$summerBPComplete[breaks$year == year & breaks$site == siteList[j]] <- TRUE
          
          plot(out,add=T,link=FALSE,lwd=2,col=2:3, lty=c(1),ylim=c(-5,25),xlim=c(0,366))  #Plot fits
          lines(out,col=2,pch=19,bottom=FALSE,lwd=2) #Plot peak point
        }
      }# End 'out' loop
    
      title(main=paste(e1$site[1],"_",year))
      dev.off()
      
    }# End numBefore/After if statement
  }# End year loop
}# End site loop

#----------------------------------------------------------------
#Merge the breakpoints and quantiles with the existing dataframe:
#----------------------------------------------------------------
#et <- merge( x=e, y=breaks, by.x=c('year','site'),by.y=c('year','site'),all.x=T)
et <- merge( x=e, y=breaks, by=c('year','site'),all.x=T)

# need to sort for plotting pred
segDF <- segDF[order(segDF$site,segDF$year,segDF$dOY),]



#Histogram of rising and falling limbs of segmented regression used in summer BP estimation.
gRisingTemp  <- ggplot( breaks) + 
  geom_histogram(aes(risingTemp ), binwidth = .01) + 
  theme(strip.text.x = element_text(size=6)) +
  scale_x_continuous(breaks = c(seq(from = -2, to = 2, by = .2)),
                     labels = c(seq(from = -2, to = 2, by = .2))) +
  ggtitle('Rising limb of seg-regression for SummerBP')

gFallingTemp <- ggplot( breaks) + 
  geom_histogram(aes(fallingTemp), binwidth = .01) + 
  theme(strip.text.x = element_text(size=6)) +
  scale_x_continuous(breaks = c(seq(from = -3, to = 2, by = .2)),
                     labels = c(seq(from = -3, to = 2, by = .2))) +
  ggtitle('Falling limb of seg-regression for SummerBP')

gHist <- arrangeGrob( gRisingTemp, gFallingTemp, ncol=1)

ggsave(plot=gHist,file=paste0(subGraphsDir,'_slopesHistogram', sourceList,'.png'),dpi=300,width=6,height=8, units='in', scale=2)

```

```{r Index abnormal plots as defined by the segmented regression.}

if (indexAbnormalSummerGraphs) {
  
  #Abnormal sites highlighted by the segmented regression:
  abnormalSiteYears
  
  setwd(paste0(subGraphsDir, '/summerBP/'))

  abnormalPlots <- paste0('summerBP_', abnormalSiteYears$site, '_' , abnormalSiteYears$year)

  sapply(c(2:length(abnormalPlots)),FUN = function(x){
    file.rename(from = paste0(abnormalPlots[x],'.png'), to = paste0(subGraphsDir,'/problemSites/', abnormalPlots[x],'.png'))})
}
```

```{r set bps}
################################################
# set segment assignments based on which BPs
# are estimated
################################################

et$segment <- NA

################################################
# Spring = YES  ||  Summer = YES || Fall = YES #
################################################

et$segment <- ifelse( et$springBPComplete & et$summerBPComplete & et$fallBPComplete, (
  
                ifelse( et$dOY <  et$springBP | et$dOY >= et$fallBP,   1,
                ifelse( et$dOY >= et$springBP & et$dOY <  et$summerBP, 2,
                ifelse( et$dOY >= et$summerBP & et$dOY <  et$fallBP,   3, et$segment ) ) ) ), et$segment )

################################################
# Spring = NO  ||  Summer = YES || Fall = YES #
################################################

et$segment <- ifelse( !et$springBPComplete & et$summerBPComplete & et$fallBPComplete, (
  
                ifelse( et$dOY >= et$fallBP,                           1,
                ifelse( et$dOY <= et$summerBP &
                        et$tempIndex > et$quantileLo & 
                        et$tempIndex < et$quantileHi,                  2,
                ifelse( et$dOY >= et$summerBP & et$dOY <  et$fallBP,   3, et$segment ) ) ) ), et$segment )

################################################
# Spring = YES  ||  Summer = YES || Fall = NO  #
################################################

et$segment <- ifelse( et$springBPComplete & et$summerBPComplete & !et$fallBPComplete, (
  
                ifelse( et$dOY <  et$springBP ,                        1,
                ifelse( et$dOY >= et$springBP & et$dOY <  et$summerBP, 2,
                ifelse( et$dOY >  et$summerBP &
                        et$tempIndex > et$quantileLo & 
                        et$tempIndex < et$quantileHi,                  3, et$segment ) ) ) ), et$segment )

################################################
# Spring = NO   ||  Summer = YES ||  Fall = NO #
################################################

et$segment <- ifelse( !et$springBPComplete & et$summerBPComplete & !et$fallBPComplete, (
  
                ifelse( et$dOY <= et$summerBP &
                          et$tempIndex > et$quantileLo & 
                          et$tempIndex < et$quantileHi,                2,
                ifelse( et$dOY >  et$summerBP &
                          et$tempIndex > et$quantileLo & 
                          et$tempIndex < et$quantileHi,                3, et$segment ) ) ), et$segment )


################################################
# Spring = YES  ||  Summer = NO  || Fall = YES #
################################################

# use mean summer segment? Yes.
et$segment <- ifelse( et$springBPComplete & !et$summerBPComplete & et$fallBPComplete, (
  
                ifelse( et$dOY <  et$springBP | et$dOY >= et$fallBP,        1,
                ifelse( et$dOY >= et$springBP & et$dOY <  et$summerBPWMean, 2,
                ifelse( et$dOY >= et$summerBPWMean & et$dOY <  et$fallBP,   3, et$segment ) ) ) ), et$segment )


################################################
# Spring = NO  ||  Summer = NO   || Fall = YES #
################################################

et$segment <- ifelse( !et$springBPComplete & !et$summerBPComplete & et$fallBPComplete, (
  
  ifelse( et$dOY >= et$fallBP,   1,
  ifelse( et$dOY <  et$fallBP & et$dOY >= et$summerBPWMean,  3, et$segment ) ) ), et$segment )
                               
################################################
# Spring = YES  ||  Summer = NO  || Fall = NO  #
################################################

et$segment <- ifelse( et$springBPComplete & !et$summerBPComplete & !et$fallBPComplete, (
  
  ifelse( et$dOY <  et$springBP,   1,
  ifelse( et$dOY >  et$springBP & et$dOY <= et$summerBPWMean,   2, et$segment ) ) ), et$segment )

################################################
# Spring = NO   ||  Summer = NO  ||  Fall = NO #
################################################

# no segments can be assigned

###########################################################################################################################

# year segment includes early winter of the next year with the current year
et$yearSegment <- et$year
et$yearSegment <- ifelse( et$dOY > et$fallBP,et$yearSegment+1,et$yearSegment )

# check order:
et <- et[order(et$count),]
```

```{r do regression for temp ~ airTemp by site, year, segment}

slopes <- ddply( et[!is.na(et$segment) & et$segment %in% 2:3,], .(year, site, segment), summarise, 
            interceptSeg = lm(temp ~ airTemp)$coefficients[1],
                slopeSeg = lm(temp ~ airTemp)$coefficients[2] )

et$slopeSeg <- NULL; et$interceptSeg <- NULL # in case we've merged before
et <- merge( x=et, y=slopes[!is.na(slopes$site),], by=c('year','site','segment'),all.x=T)
et <- et[order(et$count),]

et$slopeSegSlopeDiff <- et$slopeSeg - et$slope
```

```{r Make graphs of the results}

#Distribution of breakpoints:
boxplot(et$springBP,et$summerBP,et$fallBP,names=c("Spring","Summer","Fall"))

hist(et$springBP)
hist(et$summerBP)
hist(et$fallBP)

# Make graphs for each site
#-------------------------
if (makeSpringFallGraphs){

for (j in 1:nSites){  
  
  print(c('making graphs:',j, 'out of ',nSites))

  #Find a better fix for this case (when there is all NA in airTemp)
  temp <- et[et$site == siteList[j] & !is.na(et$temp) & !is.na(et$airTemp),]
  
if (nrow(temp) > 0 ){
  
  gBP <- ggplot( temp, aes(dOY,tempIndex)) + 
    geom_point(aes(dOY,movingSD  ),size = 0.5, colour='red' ) +
    geom_point(aes(dOY,movingMean),size = 0.5, colour='blue') +
    geom_point(size=0.5) +
    geom_hline( aes(yintercept=as.numeric(quantileLo)), colour='black') +
    geom_hline( aes(yintercept=as.numeric(quantileHi)), colour='black') +
    geom_vline( aes(xintercept=as.numeric(springBP  )), colour='green') +
    geom_vline( aes(xintercept=as.numeric(fallBP    )), colour='orange')+
    geom_vline( aes(xintercept=as.numeric(summerBP  )), colour='blue')  +
    ylim(c(-10,10))+
   #  xlim(c(80,120))+
    theme(strip.text.x = element_text(size=6))+
    ggtitle(paste(temp$site[1],sep=" ")) +
    facet_wrap(~year)
  
    #dev.off() 
    
  #gSeg <- ggplot( temp, aes( airTemp,temp, color=factor(segmentWFill) ) ) +
  if(any(!is.na(temp$segment))){    
  gSeg <- ggplot( temp, aes( airTemp,temp, color=factor(segment) ) ) +
    geom_point() +
    geom_smooth(method='lm') +  
    theme(legend.position="none") +
    theme(strip.text.x = element_text(size=6))+
  #  scale_colour_brewer(palette="Set1") +
  #    scale_colour_hue()+
  #    scale_colour_manual(values = c('blue', "green",'red','grey')) +
    facet_wrap(~year)
  
  gOut <- arrangeGrob( gBP, gSeg, ncol=1 )
  }
  
  if(all(is.na(temp$segment))) { gOut <- arrangeGrob( gBP,  ncol=1 ) }
    
    ggsave(plot=gOut,file=paste(subGraphsDir,'/springFallBP/',temp$site[1],".png",sep=''),dpi=300,width=6,height=8, units='in', scale=2)
} # nrow loop
  
} # Site loop
} # if MakeSpringFallGraphs loop

```

```{r save e}

  save(e, file=paste(dataOutDir, '/e', sourceList, '.RData',sep=''))
  save(et,segDF,slopes, file=paste(dataOutDir, '/et', sourceList, '.RData',sep=''))

```
