This script averages the local daymet climate data over the upstream catchments. 
Currently, this is only done for precipitation for the prediction catchments.

```{r Create master precipitation dataframe}

rm(list = ls())

library(reshape)

DaymetTiles <- c(11754, 11755, 11756, 11934, 11935, 11936, 12114, 12117, 12115, 12116, 12295, 12296, 12297)
Year <- 2010

setwd("C:/KPONEIL/USGS/Stream Temperature/data/temperature/fromKyle/BP_Analysis/BP_Analysis/DaymetClimateData/LocalValues")
for ( i in 1:length(DaymetTiles)){
  
  print(i)
  load(paste0('NHD_DaymetTile_' , DaymetTiles[i], '_', Year, '.RData'))
  
  FullRecord <- FullRecord[c('FEATUREID', 'dOY', 'prcp')]
  
  if( i == 1) {PRCP <- FullRecord } else ( PRCP <- rbind(PRCP, FullRecord))
  rm(FullRecord)
}

```


```{r Average upstream precipitation values}
load("C:/KPONEIL/USGS/GIS/Covariate Stats/DelineatedCatchments/DelineatedCatchments_NHDPlus_NENY.RData")
DelineatedCatchmentsMaster <- NENYDelineatedCatchments
MasterLength <- length(DelineatedCatchmentsMaster)


for ( i in 1:length(DaymetTiles)){
  
  load(paste0('NHD_DaymetTile_' , DaymetTiles[i], '_', Year, '.RData'))
  
  sites <- unique(FullRecord$FEATUREID)
  
  for( j in 1:length(sites)){
    
    print(paste0( 'Processing... Site: ', j , ' of ', length(sites),' | Tile: ', i, ' of ', length(DaymetTiles)))
    
    featureID <- sites[j]
    features <- DelineatedCatchmentsMaster[[which(sapply(c(1:MasterLength),FUN = function(x){DelineatedCatchmentsMaster[[x]][1]==featureID})==TRUE)]]
    
    Basin <- PRCP[which(PRCP$FEATUREID %in% features),]
    BasinCast <- cast(Basin, dOY~FEATUREID, value = 'prcp') 
    
    if ( length(BasinCast$dOY) > 365) {MEAN <- rowMeans(BasinCast[,-1], na.rm = T)} else(MEAN <- BasinCast[,2])    
    
    BasinMean <- data.frame(featureID, BasinCast$dOY, MEAN)
    
    names(BasinMean) <- c('FEATUREID', 'dOY', 'prcp')
    
    if( j ==1 ) { FullPrcp <- BasinMean} else ( FullPrcp <- rbind(FullPrcp, BasinMean))
  }

  FullRecord <- FullRecord[, -which(names(FullRecord) == 'prcp')]
  
  FullRecord <- merge(FullRecord, FullPrcp, by = c('FEATUREID', 'dOY'), all.x = T, sort = F)
  
  save(FullRecord, file = paste0('C:/KPONEIL/USGS/Stream Temperature/data/temperature/fromKyle/BP_Analysis/BP_Analysis/DaymetClimateData/NHD_DaymetTile_' , DaymetTiles[i], '_', Year, '.RData'))
 
}

```




##################################################################################################################################################
#                                                                                                                                                #
#                                              UNDER CONSTRUCTION                                                                                #    
#                                                                                                                                                #      
##################################################################################################################################################

```{r parallel processing version}


sites <- unique(PRCP$FEATUREID)

#DelinationStation<- function(j){
  
  for( j in 1:length(sites)){
    
    print(paste0( 'Processing... Site: ', j , ' of ', length(sites),' | Tile: ', i, ' of ', length(DaymetTiles)))
    
    featureID <- sites[j]
    features <- DelineatedCatchmentsMaster[[which(sapply(c(1:MasterLength),FUN = function(x){DelineatedCatchmentsMaster[[x]][1]==featureID})==TRUE)]]
    
    Basin <- PRCP[which(PRCP$FEATUREID %in% features),]
    BasinCast <- cast(Basin, dOY~FEATUREID, value = 'prcp') 
    
    if ( length(BasinCast$dOY) > 365) {MEAN <- rowMeans(BasinCast[,-1])} else(MEAN <- BasinCast[,2])
    #if ( length(features) > 1) {MEAN <- rowMeans(BasinCast[,-1])} else(MEAN <- BasinCast[,2])
    
    BasinMean <- data.frame(featureID, BasinCast$dOY, MEAN)
    
    names(BasinMean) <- c('FEATUREID', 'dOY', 'prcp')
    
    
    if( j ==1 ) { FullPrcp <- BasinMean} else ( FullPrcp <- rbind(FullPrcp, BasinMean))
  }

  FullRecord <- FullRecord[, -which(names(FullRecord) == 'prcp')]
  
  FullRecord <- merge(FullRecord, FullPrcp, by = c('FEATUREID', 'dOY'), all.x = T, sort = F)
  
  save(FullRecord, file = paste0('C:/KPONEIL/USGS/Stream Temperature/data/temperature/fromKyle/BP_Analysis/BP_Analysis/DaymetClimateData/UpstreamPrcp/NHD_DaymetTile_' , DaymetTiles[i], '_', Year, '.RData'))
 
}




library(parallel)

values <- 1:length(sites)

# Number of workers (R processes) to use:
numWorkers <- 8
# Set up the 'cluster':
cl <- makeCluster(numWorkers, type = "PSOCK")

# Load the necessary objects into the cluster:
clusterExport(cl, c("NewMean", "sites", "PRCP", "DelineatedCatchmentsMaster",  "MasterLength"))
clusterExport(cl, library(reshape))

# Execute the function:
start.time <- proc.time()[3]
Results    <- parSapply(cl, values, UpstreamSpatialAverage)
end.time   <- proc.time()[3]
print(paste0((end.time-start.time)/3600, " hours"))

# Shut down cluster:
stopCluster(cl)



```

  
  if (i == 1 ) {a <- FullRecord}
  if (i == 2 ) {b <- FullRecord}
  if (i == 3 ) {c <- FullRecord}
  if (i == 4 ) {d <- FullRecord}
  if (i == 5 ) {e <- FullRecord}
  if (i == 6 ) {f <- FullRecord}
  if (i == 7 )  {g <- FullRecord}
  if (i == 8 )  {h <- FullRecord}
  if (i == 9 )  {i <- FullRecord}
  if (i == 10 ) {j <- FullRecord}
  if (i == 11 ) {k <- FullRecord}
  if (i == 12 ) {l <- FullRecord}
  rm(FullRecord)
}
  