#Author:
#  Kyle O'Neil
#Created:
#  7/10/2014
#Last Updated:
#  7/10/14
#Language:
#	 R
#Description:
#  This script generates a dataframe of Daymet prediction variables for observed stream temperature sites.

#  Currently the Daymet NetCDF files cover the years 1980-2013 and the spatial range of which is 
#  VT, NH, MA, CT, RI, ME, and parts of NY.
#========================================================================================================


# Specify libraries and directories and define the projection of the spatial data:
# ================================================================================
``` {r Libraries and directories}
rm(list=ls())

#library(sp)
#library(rgdal)
#library(rgeos)
library(maptools)
#library(chron)
#library(ncdf)
#library(ggplot2)
#library(reshape2)
#library(gridExtra)

# Base directory
baseDir <- 'C:/KPONEIL/GitHub/projects/temperatureProject/'

# Daymet data directory
daymetDir <- 'F:/KPONEIL/SourceData/climate/DAYMET/unzipped/Daily/'

# dataIn directory
dataInDir <- paste0(baseDir, "dataIn/")

# Define the projection of the spatial data
proj4.NHD  <- "+proj=longlat +ellps=GRS80 +datum=NAD83 +no_defs"

# Source functions
source(paste0(baseDir, 'code/functions/temperatureModelingFunctions.R'))
source(paste0(baseDir, 'code/functions/dataIndexingFunctions.R'))

setwd(baseDir)
```

# Read in data
```{r Choose data sources}

# Select "T" or "F" for using agency data
sources <- list (
  # Northeast
  CTDEP  = CTDEP  <- F,
  MADEP  = MADEP  <- T,
  MAFW   = MAFW   <- T,
  MAUSGS = MAUSGS <- T,
  MEDMR  = MEDMR  <- F,
  MEFWS  = MEFWS  <- F,
  NHDES  = NHDES  <- F,
  NHFG   = NHFG   <- F,
  USFS   = USFS   <- F,
  VTFWS  = VTFWS  <- F,
  
  # Montana
  MTUSGSYellowstone = MTUSGSYellowstone <- F,
  MTUSGSGlacier     = MTUSGSGlacier     <- F
)

dataSources <- names(sources[(sources == T)])


# Read in covariate data using the read data function
# ---------------------------------------------------
covariateData <- readStreamTempData(timeSeries = F, covariates = T, dataSourceList = dataSources, fieldListTS = NA, fieldListCD = "ALL", directory = dataInDir )
```


```{r }

# Define the prediction record
record <- seq(from=as.Date("1980-01-01"),to=as.Date("2013-12-31"), by="day")

# Create a list of sites
sites <- unique(covariateData$site)

# Create the dataframe for the Daymet indexing functions
# ------------------------------------------------------
for( i in 1:length(sites) ){
  
  print(i)
  
  Lat <- covariateData$Latitude [covariateData$site == sites[i]]
  Lon <- covariateData$Longitude[covariateData$site == sites[i]]
  
  temp <- data.frame(site = sites[i], date = record, Latitude = Lat, Longitude = Lon)
  
  if( i == 1 ) {masterData <- temp} else( masterData <- rbind(masterData, temp))
  
}

masterData$year <- as.numeric(strftime(masterData$date, '%Y'))
masterData$dOY  <- as.numeric(strftime(masterData$date, '%j'))
```

# Index the local Daymet variables for the observed stream temperature sites:
# ===========================================================================
# This function pulls the Daymet variables for a time series at one location nearest to the site location.
# It takes the stream temperature record and the string of variables to pull from Daymet.
#     At minimum, the record needs a "site", "Latitude", "Longitude", "year", and "dOY" columns.
# It returns the original dataframe with new columns for the Daymet variables.
```{r Local Daymet}

# List the Daymet variables to index
localVariables <- c("dayl", "srad", "swe", "tmax", "tmin", "vp")

# Run the function that indexes the Daymet data
newLocal <- indexLocalDaymetVariablesForObservedSites(masterData, localVariables, daymetDir)

# Rename 'tmin' and 'tmax' to avoid confusion
names(newLocal)[names(newLocal) == 'tmax'] <- 'maxAirTemp'
names(newLocal)[names(newLocal) == 'tmin'] <- 'minAirTemp'

# Check the amount of NAs
length(which(is.na(newLocal)))/length(which(!is.na(newLocal)))*100
```

```{r Save the Local Daymet}
masterData <- newLocal[order(newLocal$site,newLocal$year,newLocal$dOY),]
save(masterData, file = paste0('C:/KPONEIL/GitHub/projects/temperatureProject/dataIn/', dataSource, '/observedStreamTempAndClimateData_', dataSource, '_NeedPrcp.RData'))


#save(masterData, file = paste0('C:/KPONEIL/GitHub/projects/temperatureProject/dataIn/daymetFullRecordObservedMASites_NeedPrcp.RData'))

```


# Index the upstream Daymet variables for the observed stream temperature sites:
# ===========================================================================
# This function calculates the spatial average of the Daymet variables within a watershed.
# It takes the following:
#    1) The stream temperature record (Site names, latitude, longitude, year, and dOY columns)
#    2) A string of variables to pull from Daymet
#    3) A list of daymet tiles covered by the watersheds
#    4) A master catchments shapefile
#    5) A dataframe of the covariates for the catchments (FEATUREIDs source)
#    6) A list of catchment delineations for the region
# It returns the original dataframe with new columns for the Daymet variables.
```{r Index the upstream Daymet variables for the observed sites}

# Load the covariate data for the featureIDs
#load(paste0(baseDir, 'temperatureProject/dataIn/', dataSource, '/covariateData_', dataSource, '.RData'))

# Daymet variables you want
upstreamVariables <- c('prcp')

# List the Daymet tiles that the watersheds may fall in
daymetTiles <- c(11754, 11755, 11756, 11934, 11935, 11936, 12114, 12115, 12116, 12117, 12295, 12296, 12297)

# Catchments shapefile
catchments <- readShapePoly ( "C:/KPONEIL/gis/nhdPlusV2/NENY_NHDCatchment.shp", proj4string=CRS(proj4.NHD))

# Delineated catchments list (variable: NENYDelineatedCatchments)
load(paste0(baseDir, 'dataIn/delineatedCatchments/DelineatedCatchments_NHDPlus_NENY.RData'))

# Run the function that indexes the Daymet data and returns a spatial average
newUpstream <- indexUpstreamDaymetVariablesForObservedSites(masterData, upstreamVariables, daymetTiles, catchments, covariateData, NENYDelineatedCatchments, daymetDirectory = daymetDir)
```


```{r Save the dataframe}
climateData <- newUpstream
save(masterData, file = paste0('C:/KPONEIL/GitHub/projects/temperatureProject/dataIn/', dataSource, '/observedStreamTempAndClimateData_', dataSource, '.RData'))

save(climateData, file = paste0('C:/KPONEIL/GitHub/projects/temperatureProject/dataIn/daymetFullRecordObservedMASites.RData'))

```