#Author:
#  Kyle O'Neil
#Created:
#  6/20/2014
#Last Updated:
#  6/25/14
#Language:
#	 R
#Description:
#  This script pairs stream temperature sites with the covariate values based on NHDPlus catchment
#  delineation. It also indexes timeseries climate data from Daymet and pairs it with the stream 
#  temperature record.

#  Currently the Daymet NetCDF files cover the years 1980-2013 and the spatial range of which is 
#  VT, NH, MA, CT, RI, ME, and parts of NY.

#  The primary input files are:
#   1) The stream temperature record
#   2) A catchments shapefile covering the contributing watersheds of this dataset
#   3) A master list of delineated catchments for the area
#   4) A master list of the local and upstream covariate values
#========================================================================================================


# Specify libraries and directories and define the projection of the spatial data:
# ================================================================================
``` {r Libraries and directories}
rm(list=ls())

library(sp)
library(rgdal)
library(rgeos)
library(maptools)
library(chron)
library(ncdf)
library(ggplot2)
library(reshape2)
library(gridExtra)

# Base directory
baseDir <- 'C:/KPONEIL/GitHub/projects/'

# Daymet data directory
daymetDir <- 'F:/KPONEIL/SourceData/climate/DAYMET/unzipped/Daily/'

# Enter the data source agency abbreviation
dataSource <- 'MEFWS'

# Define the projection of the spatial data
proj4.NHD  <- "+proj=longlat +ellps=GRS80 +datum=NAD83 +no_defs"


# Source functions
source(paste0(baseDir, 'code/functions/temperatureModelingFunctions.R'))
source(paste0(baseDir, '/code/functions/dataIndexingFunctions.R'))

setwd(baseDir)
```

# Select the agencies to use data from.
```{r Choose data source}


load('C:/KPONEIL/GitHub/projects/temperatureProject/dataIn/MADEP/covariateData_MADEP.RData')
c1 <- covariateData

load('C:/KPONEIL/GitHub/projects/temperatureProject/dataIn/MAFW/covariateData_MAFW.RData')
c2 <- covariateData

load('C:/KPONEIL/GitHub/projects/temperatureProject/dataIn/MAUSGS/covariateData_MAUSGS.RData')
c3 <- covariateData

covariateData <- rbind(c1, c2, c3)


record <- seq(from=as.Date("1980-01-01"),to=as.Date("2013-12-31"), by="day")

sites <- unique(covariateData$site)

for( i in 1:length(sites) ){
  
  Lat <- covariateData$Latitude[covariateData$]
  Lon
  
  
  data.frame(site = sites[i], date = record,  
  
  
  
  
  
  
  
  
}














# Select "T" or "F" for using agency data
sources <- list (
  # Northeast
  CTDEP  = CTDEP  <- F,
  MADEP  = MADEP  <- T,
  MAFW   = MAFW   <- T,
  MAUSGS = MAUSGS <- T,
  MEDMR  = MEDMR  <- F,
  MEFWS  = MEFWS  <- F,
  NHDES  = NHDES  <- F,
  NHFG   = NHFG   <- F,
  USFS   = USFS   <- F,
  VTFWS  = VTFWS  <- F,
  
  # Montana
  MTUSGSYellowstone = MTUSGSYellowstone <- F,
  MTUSGSGlacier     = MTUSGSGlacier     <- F
)

dataSources <- names(sources[(sources == T)])

# Enter the common fields from the temperature ("site" must be one).
tempFields <- c('site', 'year', 'dOY', 'date', 'agency', 'temp', 'airTemp')

# Enter the specific covariate fields you want to pull ("site" must be one) or for the entire file, enter "ALL"
covFields <- c('site', 'HUC4', 'HUC8', 'HUC12')

# Read in data records and join into one dataframe
# ------------------------------------------------
e <- readStreamTempData(timeSeries = T, covariates = T, dataSourceList = dataSources, fieldListTS = tempFields, fieldListCD = covFields, directory = dataInDir )

```

# Index the local Daymet variables for the observed stream temperature sites:
# ===========================================================================
# This function pulls the Daymet variables for a time series at one location nearest to the site location.
# It takes the stream temperature record and the string of variables to pull from Daymet.
#     At minimum, the record needs a unique site ID, latitude, longitude, year, and dOY columns.
# It returns the original dataframe with new columns for the Daymet variables.
```{r Local Daymet}

# List the Daymet variables to index
localVariables <- c("dayl", "srad", "swe", "tmax", "tmin", "vp")

# Run the function that indexes the Daymet data
newLocal <- indexLocalDaymetVariablesForObservedSites(masterData, localVariables)

# Rename 'tmin' and 'tmax' to avoid confusion
names(newLocal)[names(newLocal) == 'tmax'] <- 'maxAirTemp'
names(newLocal)[names(newLocal) == 'tmin'] <- 'minAirTemp'

# Check the amount of NAs
length(which(is.na(newLocal)))/length(which(!is.na(newLocal)))*100
```

```{r Save the Local Daymet}
masterData <- newLocal[order(newLocal$site,newLocal$year,newLocal$dOY),]
save(masterData, file = paste0('C:/KPONEIL/GitHub/projects/temperatureProject/dataIn/', dataSource, '/observedStreamTempAndClimateData_', dataSource, '_NeedPrcp.RData'))
```


# Index the upstream Daymet variables for the observed stream temperature sites:
# ===========================================================================
# This function calculates the spatial average of the Daymet variables within a watershed.
# It takes the following:
#    1) The stream temperature record (Site names, latitude, longitude, year, and dOY columns)
#    2) A string of variables to pull from Daymet
#    3) A list of daymet tiles covered by the watersheds
#    4) A master catchments shapefile
#    5) A dataframe of the covariates for the catchments (FEATUREIDs source)
#    6) A list of catchment delineations for the region
# It returns the original dataframe with new columns for the Daymet variables.
```{r Index the upstream Daymet variables for the observed sites}

# Load the covariate data for the featureIDs
load(paste0(baseDir, 'temperatureProject/dataIn/', dataSource, '/covariateData_', dataSource, '.RData'))

# Daymet variables you want
upstreamVariables <- c('prcp')

# List the Daymet tiles that the watersheds may fall in
daymetTiles <- c(11754, 11755, 11756, 11934, 11935, 11936, 12114, 12115, 12116, 12117, 12295, 12296, 12297)

# Run the function that indexes the Daymet data and returns a spatial average
newUpstream <- indexUpstreamDaymetVariablesForObservedSites(masterData, upstreamVariables, daymetTiles, catchments, covariateData, NENYDelineatedCatchments)
```


```{r Save the dataframe}
masterData <- newUpstream
save(masterData, file = paste0('C:/KPONEIL/GitHub/projects/temperatureProject/dataIn/', dataSource, '/observedStreamTempAndClimateData_', dataSource, '.RData'))
```