# Clean up the raw data from MADEP.
# This code reads in and formats the data.
# The code does not fill in the entire calendar year of the record with NAs as is done in previous versions.

```{r Set directories}
rm(list=ls())

library(xlsx)
library(reshape2)

baseDir <- 'C:/KPONEIL/GitHub/projects/temperatureProject/rawDataWork/'

dataSource <- 'MEFWS'

source('C:/KPONEIL/GitHub/projects/temperatureProject/code/functions/rawDataProcessingFunctions.R')

# To prevent a memory usage error, set RAM limit:
options(java.parameters = "-Xmx16000m")
```

```{r Read in and aggregate raw data}
setwd(paste0(baseDir, dataSource, '/dataIn'))

# Read the sheets of the excel file
meanT <- read.xlsx('emac_watertemps3.xlsx', sheetName='dailymean', as.data.frame=TRUE)
minT <- read.xlsx('emac_watertemps3.xlsx', sheetName='dailymin', as.data.frame=TRUE)
maxT <- read.xlsx('emac_watertemps3.xlsx', sheetName='dailymax', as.data.frame=TRUE)
siteInfo <- read.xlsx('emac_watertemps3.xlsx', sheetName='Site Info', as.data.frame=TRUE)

# Melt the dataframes
meanTMelt <- melt(meanT, id.vars="Month.Day")
minTMelt <- melt(minT, id.vars="Month.Day")
maxTMelt <- melt(maxT, id.vars="Month.Day")

# Rename
names(meanTMelt) <- c('date', 'AgencyID', 'temp')
names(minTMelt) <- c('date', 'AgencyID', 'minTemp')
names(maxTMelt) <- c('date', 'AgencyID', 'maxTemp')

# Merge them all together
merge1 <- merge(meanTMelt, minTMelt, by = c('date','AgencyID'))
record <- merge(merge1, maxTMelt, by = c('date','AgencyID'))

# Pull X's off of the siteIDs
record$AgencyID <- substr(record$AgencyID, 2, 5)

# Assign columns
record$agency <- dataSource
record$site   <- paste0(dataSource, '_', record$AgencyID)
record$year   <- as.numeric(strftime(record$date, '%Y'))
record$dOY    <- as.numeric(strftime(record$date, '%j'))

# Reorder
record <- record[order(record$site,record$year,record$dOY),]

# Pull the site that has no data:
record <- record[!record$site == "MEFWS_8",]

# Prepare the Metadata
newInfo <- siteInfo[,c('New.NSO', 'X', 'Y', 'Description')]

names(newInfo) <- c('AgencyID', 'Longitude', 'Latitude', 'description')

record <- merge(record, newInfo, by = 'AgencyID', all.x = T, all.y = F, sort = F)

# Function that trims the NAs off of the ends of the record.
# Give it a the record dataframe and the name of the column to use for trimming.

masterData <- trimNAsFromRecord(record, "temp")


```

```{r Save the data}

save(masterData, file = paste0('C:/KPONEIL/GitHub/projects/temperatureProject/dataIn/', dataSource, '/streamTempData_', dataSource, '.RData'))


```



```{r location file}

siteLocs <- unique(finalData[,c('site', 'Latitude', 'Longitude', 'agency')])

write.csv(siteLocs, file = 'C:/KPONEIL/GitHub/projects/temperatureProject/maps/siteLocations/MADEP.csv')



```




  names(dataIn) <- c('date',  'meanTemp', 'minTemp', 'maxTemp')



          as.data.frame=TRUE, header=TRUE, colClasses=NA,
          keepFormulas=FALSE, encoding="unknown")



read.xlsx(files[i], sheetIndex, sheetName=NULL, rowIndex=NULL,
          startRow=NULL, endRow=NULL, colIndex=NULL,
          as.data.frame=TRUE, header=TRUE, colClasses=NA,
          keepFormulas=FALSE, encoding="unknown", ...)