# Clean up the raw data from MADEP.
# This code reads in and formats the data.
# The code does not fill in the entire calendar year of the record with NAs as is done in previous versions.

```{r Set directories}

rm(list=ls())

library(xlsx)

baseDir <- 'C:/KPONEIL/GitHub/projects/temperatureProject/rawDataWork'

locations <- read.csv(paste0(baseDir, '/MADEP/dataIn/tblLocationDetails.csv'))

agency <- 'MADEP'


setwd(paste0(baseDir, '/MADEP/dataIn/temperatureFiles'))
files <- list.files()

# To prevent a memory usage error:

# Set RAM limit
options(java.parameters = "-Xmx16000m")

# Write a function to take out the garbage:

```

```{r Read in and aggregate raw data}

for( i in 1:length(files)){

  gc()
  
  print(i)
  
  # Read in raw data:
  dataIn <- read.xlsx(files[i], sheetName='Daily Statistics', colIndex=c(2,3,4,5), as.data.frame=TRUE)

  # Pull site ID:
  agencyID <- substr(files[i], 1, 5)

  # Store all data:
  siteData <- data.frame( site = paste0(agency, '_', agencyID), dataIn, 
                          Latitude = locations$Latitude[locations$UNIQUE_ID == agencyID], 
                          Longitude = locations$Longitude[locations$UNIQUE_ID == agencyID], 
                          year = as.numeric(strftime(dataIn$Date, '%Y')),
                          dOY = as.numeric(strftime(dataIn$Date, '%j')),
                          agency = agency,
                          AgencyID = agencyID
                          )

  # Check to make sure there are no missing records:
  full <- seq(from=as.Date(min(siteData$Date)),to=as.Date(max(siteData$Date)),by="day")
  if(nrow(siteData) < length(full) ) {print(paste0('There is a missing record on:', full[!full %in% siteData$Date]))}

  # Join sites:
  if( i == 1 ) {finalData <- siteData} else( finalData <- rbind(finalData, siteData))
  }
```

```{r Post-processing}

# Rename date field
finalData$date <- finalData$Date

# Put entries into numeric format (NAs will be introduced by coercion)
finalData$meanTemp <- as.numeric(paste(finalData$Temperature.......................Daily.Mean))
finalData$minTemp <-  as.numeric(paste(finalData$Temperature.......................Daily.Min ))
finalData$maxTemp <-  as.numeric(paste(finalData$Temperature.......................Daily.Max ))

finalData <- finalData[, - which(names(finalData) %in% c('Temperature.......................Daily.Mean', 
                                                         'Temperature.......................Daily.Min', 
                                                         'Temperature.......................Daily.Max',
                                                         'Date'))]

# Fix some factor/character BS:
finalData$site <- paste(finalData$site)
finalData$agency <- paste(finalData$agency)
finalData$AgencyID <- paste(finalData$AgencyID)




# Save the data:
save(finalData, file = paste0(baseDir,'/MADEP/dataOut/streamTempDataOnlyMADEP.RData'))


masterData <- finalData

save(masterData, file = paste0('C:/KPONEIL/GitHub/projects/temperatureProject/dataIn/MADEP/streamTempData_MADEP.RData'))


```



```{r location file}

siteLocs <- unique(finalData[,c('site', 'Latitude', 'Longitude', 'agency')])

write.csv(siteLocs, file = 'C:/KPONEIL/GitHub/projects/temperatureProject/maps/siteLocations/MADEP.csv')



```




  names(dataIn) <- c('date',  'meanTemp', 'minTemp', 'maxTemp')



          as.data.frame=TRUE, header=TRUE, colClasses=NA,
          keepFormulas=FALSE, encoding="unknown")



read.xlsx(files[i], sheetIndex, sheetName=NULL, rowIndex=NULL,
          startRow=NULL, endRow=NULL, colIndex=NULL,
          as.data.frame=TRUE, header=TRUE, colClasses=NA,
          keepFormulas=FALSE, encoding="unknown", ...)